### 2013-09-12 icml/proof.tex
small edits to proof of lemma and updated comment to reflect changes

\documentclass{article}
\usepackage{import}
\input{latex-defs.tex}

\begin{document}

Throughout this section, assume that $\alpha_1,\ldots,\alpha_T$ is 
a sequence of real numbers and that $A_t = \sum_{s=1}^t \alpha_s$. 
We further require that $A_t \geq 0$ for all $t$.

Also recall that the Bregman divergence is defined by $D_f(x_2 \| x_1) \eqdef f(x_2) - \langle \nabla f(x_1), x_2-x_1 \rangle - f(x_1)$.

A key observation in the following proof is that, if $x_0 = \arg\min f(x) + \langle z, x \rangle$, then $\nabla f(x_0) = -z$.

\NA{JNS: In the following proof, (2) is by the update formula for $x_{t+1}$ together with the preceding observation.  (3) is applying (2). (4) and (5) are re-arranging the sum. (6) is definition of Bregman divergence. (7) is the observation again. 
(8) is the definition of $x_{T+1}$.}

\begin{lemma}
\label{lem:ftl}
Let $z_1,\ldots,z_T$ be vectors and let $f(x)$ be a strictly convex 
function. Define $\hat{z}_t$ to be $\frac{1}{A_t} \sum_{s=1}^t \alpha_s z_s$.

Let $x_1,\ldots,x_T$ be chosen via $x_{t+1} = \arg\min_{x} f(x) + \langle \hat{z}_t, x\rangle$. 
Then for any $x^*$ we have
\begin{equation}
\frac{1}{A_T} \sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\} \leq f(x^*) + \langle \hat{z}_t, x^* \rangle + \frac{1}{A_T} \sum_{t=1}^T A_t D_{f}(x_t \| x_{t+1}). 
\end{equation}
\end{lemma}
\begin{proof}
Note that 
\begin{align}
\alpha_{t}z_{t} = A_{t}\hat z_{t} - A_{t-1}\hat z_{t-1} = - A_{t}\nabla f(x_{t+1}) + A_{t-1} \nabla f(x_{t}) 
\end{align}
so we have
\begin{align}
\lefteqn{\sum_{t=1}^T \{\alpha_t(f(x_t) + \langle z_t, x_t \rangle)\}} \\
 &= \sum_{t=1}^T \{\alpha_t f(x_t) + \langle A_{t-1} \nabla f(x_t) - A_t \nabla f(x_{t+1}), x_t \rangle\} \\
 &= \sum_{t=1}^T \{\alpha_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle\} - A_{T}\langle \nabla f(x_{T+1}), x_{T+1} \rangle \\
 &= \sum_{t=1}^T \{A_t f(x_t) - \langle A_{t} \nabla f(x_{t+1}), x_t-x_{t+1} \rangle - A_t f(x_{t+1})\}  \\
 &\quad+ A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \nonumber \\
 &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) - \langle \nabla f(x_{T+1}), x_{T+1} \rangle) \\
 &= \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x_{T+1}) + \langle \hat{z}_T, x_{T+1} \rangle) \\
 &\leq \sum_{t=1}^T \{A_tD_f(x_t \| x_{t+1})\} + A_T(f(x^*) + \langle \hat{z}_T, x^* \rangle). \\
\end{align}
Dividing both sides by $A_T$ completes the proof.
\end{proof}
We also note that $D_f(x_t \| x_{t+1}) = D_{f^*}(\hat{z}_{t+1} \| z_t)$, where $f^*(z) = \sup_x \langle z,x\rangle - f(x)$. 
This form of the bound will often be more useful to us.
\end{document}