### 2013-10-01 icml/lp.tex
added L^p argument

\documentclass[reqno,oneside,a4paper]{amsart}
%\usepackage{jhh-misc}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{natbib}
\usepackage{subfig}
\usepackage{xspace}
\usepackage[normalem]{ulem}

\input{latex-defs.tex}
% inner product command
\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
\def\[#1\]{\begin{align}#1\end{align}}

\begin{document}

\title{Herding convergence for $L^p$ norm} 
\author{Jacob Steinhardt}
\date{\today}

\maketitle

Suppose that we have the loss function $L(u) = \frac{1}{p} \|u-u^*\|_p^p$, where $p > 1$. 
We will show that a modified herding algorithm approaches $u^*$ at a rate of $O(1/T)$.

First, note that $L(u) = \sup_{\theta \in \bR^d} \ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q$, 
for $\frac{1}{p} + \frac{1}{q} = 1$ (see Lemma 1).

Now, if we perform herding updates with $R(\theta) = \ip{u^*}{\theta} + \frac{1}{q}\|\theta\|_q^q$, we 
get a convergence rate of
\[ \sum_{t=1}^T \alpha_t D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t). \]
The key part of our argument is to analyze $D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t)$. For convenience, 
let $x = \hat{u}_{t+1}-u^*$, $y = \hat{u}_t-u^*$, and $z = u_{t+1}-u^*$.
First, note that 
\begin{align*}
D_{R^*}(\hat{u}_{t+1} \| \hat{u}_t) &= \frac{1}{p} \sum_{i=1}^d x_i^p-py_i^{p-1}(x_i-y_i)-y_i^p \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)\left(\sum_{r=0}^{p-2} y_i^rx_i^{p-r-1}-y_i^{p-1}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{r=0}^{p-2} \sum_{s=0}^{p-r-2} y_i^ry_i^sx_i^{p-(r+s)-2}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{r+s=0}^{p-2} y_i^{r+s}x_i^{p-(r+s)-2}\right) \\
 &= \frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right).
\end{align*}
Now as a side calculation, we have $y_i = \frac{1}{A_t}\sum_{s \leq t} \alpha_s (u_{i,s}-u^*)$ 
and $x_i = \frac{1}{A_{t+1}} \sum_{s \leq t+1} \alpha_s (u_{i,s}-u^*)$. If we simplify we get
$(x_i-y_i)^2 = \frac{\alpha_{t+1}^2}{A_{t+1}^2} (y_i-z_i)^2$. Therefore, we have
\begin{align*}
\lefteqn{\frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right)} \\
 &\leq \frac{\alpha_{t+1}^2}{p A_{t+1}^2} \sum_{t=0}^{p-2} \sum_{i=1}^d 2(t+1)(y_i^2+z_i^2)y_i^{t}x_i^{p-t-2}.
\end{align*}
This latter sum has $4\binom{p}{2}$ terms in it, each of the form $\sum_{i=1}^d x_i^ay_i^bz_i^c$ where $a+b+c=p$. 
But we have $x_i^ay_i^bz_i^c \leq \frac{ax_i^p+by_i^p+cz_i^p}{p}$, so $\sum_{i=1}^d x_i^ay_i^bz_i^c \leq \max(\|x\|_p^p, \|y\|_p^p, \|z\|_p^p)$.
Therefore, we obtain the bound
\begin{align*}
\lefteqn{\frac{1}{p} \sum_{i=1}^d (x_i-y_i)^2 \left(\sum_{t=0}^{p-2} (t+1)y_i^{t}x_i^{p-t-2}\right)} \\
 &\leq \frac{4\alpha_{t+1}^2}{p A_{t+1}^2}\binom{p}{2} \sup_{u \in U} \|u-u^*\|_p^p \\
 &= \frac{2\alpha_{t+1}^2(p-1)(2r_p)^p}{A_{t+1}^2}.
\end{align*}
This yields an overall bound of $\frac{8(p-1)(2r_p)^p}{T}$ for the step size $\alpha_t = t$.
\begin{remark}
We can probably shave the $2r_p$ down to just $r_p$ if we're more careful with this argument.
\end{remark}
\begin{remark}
This appears to work for any $p > 1$.
\end{remark}

\section{Proof of Lemma 1}
We want to compute $\sup_{\theta \in \bR^d} \ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q$. To do this, 
we will take the derivative with respect to $\theta_i$ and set it to $0$, which yields:
\[ u_i - u_i^* = \theta_i^{q-1}. \]
Therefore, we have
\[ \theta_i = \left(u_i-u_i^*\right)^{\frac{1}{q-1}}. \]
Plugging back in, we get
\begin{align*}
\ip{u-u^*}{\theta} - \frac{1}{q} \|\theta\|_q^q &= \sum_{i=1}^d (u_i-u_i^*)^{\frac{q}{q-1}} \cdot \left(1-\frac{1}{q}\right) \\
 &= \frac{1}{p} \|u-u^*\|_p^p,
\end{align*}
as was to be shown.

\end{document}