### 2013-10-03 icml/herding-connection.tex
split into subfiles

\documentclass[paper.tex]{subfiles}
\begin{document}
\section{Application to Herding}
\label{sec:herding}

\NA{JHH: Need to specify the spaces we are working with here.}
Suppose we are given a family of features $\phi$ together 
with a known value $\bar{\phi}$ for $\bE_{u}[\phi]$. We 
would like to construct a distribution that approximately 
matches this distribution, i.e.~one for which 
$\bE_{\hat{u}}[\phi] \approx \bar{\phi}$. A natural way 
to do this is by using boosted mirror descent to minimize the 
maximum mean discrepancy relative to $\phi$; 
in other words, to minimize
\[ L(u) = \max_{i \in \{1,\dots,d\}} |\bE_{u}[\phi_i]-\bar{\phi}_i|. \]
If we wish to write this more similarly to the original problem 
formulation, we can write
\[ L(u) = \sup_{\|\theta\|_1 \leq 1} \theta^T(\bE_{u}[\phi]-\bar{\phi}). \]
One issue is that $\|\theta\|_1$ is not strongly convex, so we relax $\|\theta\|_1$ 
to $\|\theta\|_2$ and write our loss function in the Lagrangian form:
\[ L(u, \theta) = \theta^T\bE_{u}[\phi] - \theta^T\bar{\phi} - \frac{1}{2}\|\theta\|_2^2. \]
Intriguingly, Algorithm 1 and Algorithm 2 are identical to each other for this choice of $L$ 
(up to a change of index). In this case, $u_t = \delta_{x_t}$ for some $x_t$, and 
the updates are (for Algorithm 2):
\begin{itemize}
\item $x_t \in \arg\min_{x} \theta_t^T\phi(x)$
\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x_{s}) - \bar \phi$.
%\item $\theta_{t+1} = \frac{1}{A_t} \sum_{s=1}^t \alpha_s \phi(x)$.
\end{itemize}
These are the same as the standard herding updates when 
$\alpha_t = 1$. Note that $R$ is strongly convex with respect to 
$\|\cdot\|_2$, which is its own dual. Therefore, if we let
$r = \sup_{x} \|\phi(x)\|_2$, then Corollary~\ref{cor:method-2} gives us
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T \delta_{x_t}, \theta\right) &\leq \inf_{u^*} \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
 &= \sup_{\theta} \inf_{u^*} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
 &= \frac{2r^2\log(T+1)}{T}.
\end{align*}
This is a slightly weaker version of the typical herding bound on MMD.
Note that if we use $\alpha_t = t$, which corresponds to using a weighted 
average, then we obtain a stronger bound of $\frac{8r^2}{T}$ instead.

\subsection{Herding in Infinite Dimensions}
\label{sec:infinite-case}

\subsection{Herding and Maximum Entropy}
\label{sec:max-ent}

Bach et al.~have observed that herding appears to be approximately building 
samples from the maximum entropy distribution for a given collection of moment 
constraints. In this section we will make this connection more explicit 
as well as draw other relationships between herding and maximum entropy.

Suppose that instead of setting $h(u) = 0$ we set $h(u)$ to a small multiple of 
the negative entropy: $h(u) = \eta \bE_{u}[\log u(x)]$. We will still use 
$R(\theta) = \frac{1}{2}\|\theta\|_2^2$. In this case, Algorithm 2 gives the updates
\begin{itemize}
\item $u_t(x) \propto \exp\left(-\frac{1}{\eta}\theta_t^T\phi(x)\right)$
\item $\theta_{t+1} = \bE_{\hat{u}_t}[\phi(x)] - \bar\phi$.
\end{itemize}
Corollary~\ref{cor:method-2} implies that for all $u^{*}$,
\begin{align*}
\sup_{\theta} L\left(\frac{1}{T} \sum_{t=1}^T u_t, \theta\right) &\leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2\log(T+1)}{T} \\
 &= -\eta H(u^*) + \frac{1}{2}\|\bE_{u^*}[\phi(x)]-\bar{\phi}\|_2^2 \\ &\phantom{=} + \frac{8r^2}{T}.
\end{align*}
In particular, let $u^*$ be the maximum entropy distribution satisfying $\bE_{u^*}[\phi(x)] = \bar{\phi}$, 
and let $H^*$ be the corresponding entropy. Then we have 
\[ -\eta H(\hat{u}) + \|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq -\eta H^* + \frac{8r^2}{T}. \]
Since $\|\cdot\|_2^2 \geq 0$, this implies that 
\begin{equation}
H(\hat{u}) \geq H^* - \frac{8r^2}{\eta T},
\end{equation}
and so asymptotically $\hat{u}$ has at least 
as much entropy as $u^*$.

Now, suppose that our state space has some finite size $S$, so 
that no distribution has entropy greater than $\log(S)$. Then 
we conclude that
\begin{equation}
\|\bE_{\hat{u}}[\phi(x)]-\bar{\phi}\|_2^2 \leq \eta[\log(S)-H^*] + \frac{8r^2}{T}.
\end{equation}

If we let $\eta \to 0$ and $\eta T \to \infty$, then we obtain 
a distribution $\hat{u}$ whose moments are arbitrarily close to 
the moments of $u^*$, and whose entropy is almost as large as 
(or larger than) the entropy of $u^*$.

\end{document}