### 2013-10-03 icml/paper.tex

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014}
\input{latex-defs.tex}

\icmltitlerunning{Boosted Mirror Descent}

\begin{document}

\twocolumn[
\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}

\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
\icmladdress{MIT CSAIL,
             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}

\vskip 0.3in
]

\begin{abstract} 
TODO
\end{abstract}

\subfile{introduction.tex}
\subfile{herding-connection.tex}
\subfile{convergence-proofs.tex}
%\subfile{lower-bounds.tex}
%\subfile{chen-proofs.tex}
%\subfile{conclusion.tex}

%\bibliography{example_paper}
%\bibliographystyle{icml2013}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.