### 2013-10-04 icml/paper.tex

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014}
\input{latex-defs.tex}

\icmltitlerunning{Boosted Mirror Descent}

\begin{document}

\twocolumn[
\icmltitle{Boosted Mirror Descent, Frank-Wolfe, and Herding}

\icmlauthor{Jonathan Huggins \NA{add footnote on authorship}}{jhuggins@mit.edu}
\icmladdress{MIT CSAIL,
             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}

\vskip 0.3in
]

\begin{abstract} 
TODO
\end{abstract}

\subfile{introduction.tex}
\subfile{herding-connection.tex}
\subfile{convergence-proofs.tex}
%\subfile{lower-bounds.tex}
%\subfile{chen-proofs.tex}
%\subfile{conclusion.tex}

\bibliography{herding.bib}
\bibliographystyle{icml2014}

\end{document}