### 2013-10-04 icml/lower-bounds.tex

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{Lower Bounds on Convergence Rates}
\label{sec:lower-bounds}

It has been an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case. We now show that this is not in general possible. In particular, the upper bounds in the previous section for herding and herding-like algorithms have matching lower bounds for at least some combinations of distributions and feature functions.

For the matching $O(1/T)$ lower bound we analyze sparse update methods. Formally, define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
\]
Sparse update methods can be viewed as characterizing ``herding-like'' algorithms, including the generalized herding and \cgd with line search. These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

The bound setup is as follows. Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero. Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$. Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$. Without loss of generality we can assume that $\alpha$ is monotonically increasing. We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

Since we are only considering sparse update methods, the $t$-th estimator will have at most $t$ non-zero entries, and so 
\[
\|x-x^{*}\|_2^2 \ge c^{2}\sum_{n > t} 1/\alpha(n)^{2} = \Theta(1/\beta(t)) \label{eq:convergence-lb}
\]
for some (monotonically increasing) superlinear function $\beta$. But for any such $\beta$ we can choose some $\alpha$ to obtain \eqref{eq:convergence-lb}. In particular, let $1/\gamma(t) = \frac{\dee}{\dee t}\frac{1}{\beta(t)}$, where $\gamma(t) \in \omega(t^{2})$. Then we can choose $\alpha = \sqrt \gamma \in \omega(t)$ to satisfy \eqref{eq:convergence-lb}.

Therefore, any sparse update method must have convergence rate no faster than $\Theta(1/\beta(t))$ for all superlinear $\beta$, and hence no faster than $\Theta(1/t)$.

\end{document}