### 2013-10-04 icml/conclusion.tex
more small fixes; tried to improve l.b. section

\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Conclusion} 
\label{sec:conclusion}

\NA{Write me!}

\end{document}