### 2013-10-04 icml/supplementary-material.tex

\documentclass[reqno,oneside,a4paper]{amsart}
%\usepackage{jhh-misc}
\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{framed}
\usepackage{natbib}
\usepackage{subfig}
\usepackage{xspace}
\usepackage[normalem]{ulem}

\input{latex-defs.tex}

\begin{document}

\title{Supplementary Material} 
\author{}
\date{}

\maketitle

\appendix

\section{CGD, MD, and Boosted MD}

The \bmd algorithms are very closely related to \cgd and \md. \dual and \cgd are related as follows. \dual minimizes $g + R^{*}$ while \cgd minimizes $h^{*} + f^{*}$. Make the identifications $h^{*} = R^{*}$ and $f^{*} = g$ and note the conditions on the functions being optimized are equivalent. Then with $\rho_{t} = \alpha_{t+1}/A_{t+1}$, we can rewrite the \cgd updates as 
\[
x_{t} &= \partial R^{*}(-\bar y_{t})  \\
y_{t} &= \partial g^{*}(x_{t}) \\
\bar y_{t+1} &= A_{t+1}^{-1}\sum_{s \le t+1} \alpha_{t} y_{t},
\]
which is identical to the \dual if we make the substitutions $x_{t} \to -\theta_{t}$ and $y_{t} \to -u_{t}$.

We similarly relate \primal and \md. \primal minimizes $g + R^{*}$ while \md minimizes $h + f$. Make the identifications $h = g$ and $f = R^{*}$ and note the conditions of $h$ and $g$ are the same. Letting $\rho_{t+1} = \alpha_{t}/A_{t}$, we can rewrite the \md updates as 
\[
y_{t} &= \partial R^{*}(x_{t}) \\
\bar y_{t} &= \partial g^{*}((1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}) \\
x_{t+1} &= \partial g^{*}(-\bar y_{t}), 
\]
where $\bar y_{t}$ is the same as above. Making the substitutions $y_{t} \to \theta_{t}$ and $x_{t} \to u_{t}$ completes the identification of the two methods.

Finally, note that if we write $\alpha_{t} = \rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$, we have $A_{t} = \prod_{s=1}^{t} (1 - \rho_{s})$ and hence $\alpha_{t} / A_{t} = \rho_{t}$. Hence, up to a change in index, we have a bijection between the $\rho_{t}$ and the $\alpha_{t}$.

\section{Proof Lemma used in $q$-Herding Convergence Proof}

Two simple equalities were used in the proof of the convergence rate of $q$-herding. We now provide proofs. Using the fact that $x^{n} - y^{n} = (x - y)\sum_{i=0}^{n-1}x^{i-1}y^{n-i-1}$, for the first inequality we have
\(
x^{p} - py^{p-1}x + (p-1)y^{p}
&= x^{p} - y^{p} - py^{p-1}x + py^{p} \\
&= (x - y)\sum_{r=0}^{p-2} (y^{r}x^{p-r-1} - y^{p-1}) \\
&= (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}x^{p-r-s-2}y^{r+s} \\
&\le (x - y)^{2}\sum_{r=0}^{p-2}\sum_{s=0}^{p-r-2}\max(x, y)^{p-2} \\
&= {p \choose 2}(x - y)^{2}\max(x, y)^{p-2}.
\)
For the second inequality we have
\(
(y-z)^{2}\max(x,y)^{p-2} 
&\le 4\max(y,z)^{2}\max(x,y)^{p-2} \\
&\le 4\max(x,y,z)^{p}.
\)

\end{document}