### 2013-10-04 icml/chen-proofs.tex
added chen proofs

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{An Upper Bound in Infinite Dimensions}
\label{sec:chen-proofs}

An intriguing aspect of classical herding is that it is known to have a convergence rate of 
$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010b}. In this section, 
we will need to assume that $h(x)$ is the indicator function for $U$ (i.e. $0$ on $U$ and 
$\infty$ outside of it) and that $R(\theta) = \ip{u_0}{\theta} + R_0(\theta)$, where 
$R_0$ exhibits \emph{quadratic scaling}: $R_0(s\theta) = s^2R_0(\theta)$. Note that 
this implies that $R_0^*(u-u_0) = R^*(u)$.
For simplicity 
we will focus on the case $\alpha_t = 1$. Then we have 
the following bound:
\begin{theorem}
TODO
\end{theorem}
\begin{proof}
The key quantity we will analyze is $(t-1)^2R(\theta_t)$. Note that we have
\begin{align*}
\lefteqn{(t-1)^2R(\theta_t) - t^2R(\theta_{t+1})} \\
 &= (t-1)^2R^*(\hat{u}_{t-1})-t^2R^*(\hat{u}_t) \\
 &= (t-1)^2R_0^*(\hat{u}_{t-1})-t^2R_0^*(\hat{u}_t-z) \\
 &= (t-1)^2\left[R^*(\hat{u}_{t-1})-R_0^*((t/(t-1))(\hat{u}_t-z))\right] \\
 &= (t-1)^2\left[R^*(\hat{u}_{t-1})-R^*\left(\frac{t\hat{u}_t-z}{t-1}\right)\right] \\
 &= (t-1)^2\left[\ip{\frac{\partial R^*}{\partial \hat{u}_{t-1}}}{\hat{u}_{t-1}-\frac{t\hat{u}_t-z}{t-1}}-D_{R^*}\left(\frac{t\hat{u}_t-z}{t-1} \middle\| \hat{u}_{t-1}\right)\right] \\
 &\geq (t-1)^2\left[\ip{\theta_t}{-\frac{1}{t-1}(u_t-z)}-\frac{1}{2}\left\|\frac{u_t-z}{t-1}\right\|_*^2\right] \\
 &= -(t-1)\ip{\theta_t}{u_t-z}-\frac{1}{2}\|u_t-z\|_*^2 \\
 &= -(t-1)\left[\inf_{u \in U} \ip{\theta_t}{u} - \ip{\theta_t}{z}\right]-\frac{1}{2}\|u_t-z\|_*^2 \\
 &\geq -(t-1)\left[\sup_{\theta} \inf_{u \in U} \ip{\theta}{u-z}\right]-2r_*^2.
\end{align*}

\end{proof}

\end{document}