### 2013-12-04 nips-poster/nips-poster.tex
first pass at structuring the NIPS poster

\documentclass[final]{beamer}
\mode<presentation>
{
 %\usetheme{Berlin}
 % \usetheme{Aachen}
%  \usetheme{Oldi6}
%  \usetheme{I6td}
 \usetheme{I6dv}
%  \usetheme{I6pd}
%  \usetheme{I6pd2}
}

\usepackage{bm}

\usepackage{times}
\usepackage{amsmath,amssymb,amsthm}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}

\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{natbib}

\input{latex-defs.tex}

%\usepackage[orientation=landscape,size=a0,scale=1.4]{beamerposter}
\usepackage[orientation=landscape,size=custom,width=121.92,height=91.44,scale=1.4]{beamerposter} % 48in x 36in

% Display a grid to help align images
%\beamertemplategridbackground[1cm]

\title%[] %(optional, use only with long paper titles)
{A Greedy Framework for First-Order Optimization}

%\subtitle {Include Only If Paper Has a Subtitle}

\author%[Author, Another] % (optional, use only with lots of authors)
{Jonathan Huggins$^{*1}$, Jacob Steinhardt$^{*2}$}

\institute[] % (optional, but mostly needed)
{$^{*}$Both authors contributed equally to this work. $^1$Massachusetts Institute of Technology. $^{2}$Stanford University.}

\date{December 2013}

\begin{document}

\begin{frame}{} 
%\titlepage  
%\vfill
\begin{columns}
\begin{column}{0.48\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 1		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Motivation}
JS
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END MOTIVATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Being (Too) Greedy}
JHH
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END TOO GREEDY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Being Greedy Enough: Boosted Mirror Descent}
JHH

This block would contain the statement of the algorithms and the geometric intuition 
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ALGORITHM BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{column}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% COLUMN 2		 						%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{column}{0.48\linewidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN APPLICATIONS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Special Cases and Applications}
JS?
\begin{columns}[t]
\begin{column}{0.45\linewidth}

\begin{block}{Subgradient descent}

\end{block}

\begin{block}{Conditional gradient descent}
\end{block}

\begin{block}{Thresholded gradient}
\end{block}

\end{column}
\begin{column}{0.45\linewidth}
\begin{block}{AHK low-rank SDP solver}
\end{block}

\begin{block}{$q$-herding}
\end{block}

\end{column}
\end{columns}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END APPLICATIONS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\large Convergence of Boosted Mirror Descent}
\begin{columns}[t]
\begin{column}{0.45\linewidth}
\begin{theorem}
Consider the primal (respectively the dual) algorithm. Suppose that $h$ ($R$) is strongly convex with respect to a norm $\|\cdot\|$ 
and let $r = \sup_{\theta} \|\theta\|_{*}$ ($r = \sup_{u} \|u\|_{*}$). Then
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2}{A_T} \sum_{t=1}^T \frac{\alpha_{t+1}^2A_t}{A_{t+1}^2}. \]
\end{theorem}
\end{column}
\begin{column}{0.45\linewidth}
\begin{corollary} 
Under the hypotheses of the Theorem, for $\alpha_{t} = 1$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{2r^2 (\log (T) + 1)}{T}. \]
and for $\alpha_t = t$ we have
\[ \sup_{\theta} L(\hat{u}, \theta) \leq \sup_{\theta} L(u^*, \theta) + \frac{8r^2}{T}. \]
\end{corollary}
\end{column}
\end{columns}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END CONVERGENCE BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BEGIN AHK APPLICATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{block}{\large Application: Solving SDPs}
%
%
%\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% END AHK APPLICATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% BEGIN Q-HERDING APPLICATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{block}{\large Application: $q$-herding}
%
%
%\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% END Q-HERDING APPLICATION BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\small Acknowledgements}
{\footnotesize Thanks to Cameron Freer, Simon Lacoste-Julien, Martin Jaggi, Percy Liang, Arun Chaganty, and Sida Wang 
for helpful discussions. Thanks also to the anonymous reviewer for suggesting improvements 
to the paper. JS was supported by the Hertz Foundation. JHH was supported by the U.S. Government under FA9550-11-C-0028 and awarded by the Department of Defense, Air Force Office of Scientific Research, National Defense Science and Engineering Graduate (NDSEG) Fellowship, 32 CFR 168a.}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END ACKNOWLEDGMENTS BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% BEGIN BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{block}{\small References}
\begin{footnotesize}
\bibliographystyle{plainnat}
\bibliography{herding}
\end{footnotesize}
\end{block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% END BIBLIOGRAPHY BLOCK %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{column}

\end{columns}
\end{frame}
\end{document}