### 2014-01-01 lp/framework.tex
added online learning framework

\documentclass[11pt]{article}
\usepackage{import}
\input{../icml/latex-defs.tex}
\usepackage{fullpage}

\title{Online Learning Framework}
\author{Jacob Steinhardt}
\begin{document}
\maketitle
We are interested in the following learning problem: our experts 
consist of vectors $w \in \sC_P$, where $\sC_P$ is some convex 
set; similarly, our adversary can play any move $z_t$ with 
$z_t \in \sC_D$. A typical example is $\sC_P$ is the $n$-dimensional 
simplex and $\sC_D$ is the unit $l^{\infty}$ ball. Our end goal is 
to obtain a regret bound of the following type:
\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\rho}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
If we let $\sigma = \sup_{w \in \sC_P, z \in \sC_D} |w|^T|z|$, then the regret is bounded by
$\frac{\rho}{\epsilon} + \epsilon \sigma T$, which is equal to $2\sqrt{\rho \sigma T}$ for 
$\epsilon = \sqrt{\frac{\rho}{\sigma T}}$. This is the type of regret bound that is typical 
in most online learning settings. However, our bound can lead to stronger results in some 
situations.

We propose the following general style of algorithm for obtaining such 
bounds. The key innovation is an auxiliary variable $x_t$ that lets us 
relate $w_t^Tz_t$ and $w^Tz_t$. The updates are as follows:
\begin{align*}
x_{t+1} &= f(x_t, \eta z_t) \\
w_{t+1} &= \argmax_{w \in \sC_P} w^Tx_{t+1} - \psi(w) \\
\Phi_{t+1} &= w_{t+1}^Tx_{t+1} - \psi(w_{t+1})  \\ &= \psi^*(x_{t+1}).
\end{align*}

To see why these updates are a good idea, we will make a ``physicist's approximation'' 
and suppose that $f(x_t, \eta z_t) \approx x_t - \eta z_t$. Then we would have
\[ x_{T+1} \approx x_1 - \eta \sum_{t=1}^T z_t \] and thus, by definition, 
\[ \Phi_{T+1} \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t. \]
On the other hand, 
\begin{align*}
\Phi_{t+1} &\approx \Phi_t + w_t^T(x_{t+1} - x_t) \\
           &= \Phi_t - \eta w_t^T z_t,
\end{align*}
so we have 
\[ \Phi_{T+1} \approx \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
Combining this with the preceding bound on $\Phi_{T+1}$ yields
$\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t \gtrsim w^Tx_1 - \psi(w) - \eta \sum_{t=1}^T w^Tz_t$, 
or
\[ \sum_{t=1}^T w_t^Tz_t \lesssim \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \sum_{t=1}^T w^Tz_t, \]
which gives us an even better regret bound than we had sought (the $\sum_{t=1}^T |w|^T|z_t|$ term 
will show up when we track the errors made in each of the approximations).

\paragraph{Formal statement.} To turn these approximations into a rigorous bound, we will 
need $f$ and $\psi$ to satisfy the following properties:
\begin{itemize}
\item Convexity of $\psi^* \circ f$: $\psi^*(f(x,\eta z)) \leq \psi^*(x) - \eta w_x^Tz$, where $w_x = \partial \psi^*(x)$.
\item Smoothness of $f$: $w^T(x-\eta z - f(x,\eta z)) \leq \eta^2 \alpha |w|^T|z|$ for all $w$.
\end{itemize}
(For some intuition, any $f$ satisfying the two properties above 
must satisfy $f(x,0) = x$ and $f(x,\eta z) = x-\eta z + O(\eta^2)$.)
\begin{theorem}
\label{thm:main}
Suppose that $\psi^*$ and $f$ satisfy the two properties above. Then, 
for $\eta = \frac{\epsilon}{\alpha}$, we have the bound
\[ \sum_{t=1}^T w_t^Tz_t \leq \alpha \frac{\Phi_1 + \psi(w) - w^Tx_1}{\epsilon} + \epsilon \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t. \]
\end{theorem}
\begin{proof}
By the first condition, we have
\begin{align*}
\Phi^{t+1} &= \psi^*(x_{t+1}) \\
 &= \psi^*(f(x_t, \eta z_t)) \\
 &\leq \psi^*(x_t) - \eta \left(\frac{\partial \psi^*}{\partial x_t}\right)^T z_t \\
 &= \Phi_t - \eta w_t^T z_t,
\end{align*}
which by induction implies that
\[ \Phi_{T+1} \leq \Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t. \]
By the second condition, we have
\begin{align*}
w^Tx_{t+1} &= w^Tf(x, \eta z) \\
 &\geq w^Tx_t - \eta w^Tz_t - \eta^2 \alpha |w|^T|z_t|,
\end{align*}
which by induction implies that
\[ w^Tx_{T+1} \geq w^Tx_1 - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t. \]
Combining this with the first inequality yields
\begin{align*}
\Phi_1 - \eta \sum_{t=1}^T w_t^Tz_t &\geq \Phi_{T+1} \\
 &\geq w^Tx_{T+1} - \psi(w) \\
 &\geq w^Tx_1 - \psi(w) - \eta^2\alpha \sum_{t=1}^T |w|^T|z_t| - \eta \sum_{t=1}^T w^Tz_t.
\end{align*}
Re-arranging yields
\[ \sum_{t=1}^T w_t^Tz_t \leq \frac{\Phi_1 + \psi(w) - w^Tx_1}{\eta} + \eta\alpha \sum_{t=1}^T |w|^T|z_t| + \sum_{t=1}^T w^Tz_t, \]
which gives the desired result.
\end{proof}
\section{Examples}
\paragraph{Multiplicative updates.} If we let $\sC_P$ be the simplex and $\sC_D$ be the 
$l^{\infty}$ ball of radius $r$, then a standard algorithm for minimizing regret is the 
\emph{multiplicative weights update method}. We recover this by letting $f(x,z) = x + \log(1-z)$ 
and $\psi(w) = \sum_{i=1}^n w_i\log w_i$. Then $\psi^*(x) = \log\left(\sum_{i=1}^n \exp(x_i)\right)$, 
and $w_{t+1,i} = \exp(x_{t+1,i})/\sum_{j=1}^n \exp(x_{t+1,j})$.
\begin{align*}
\psi^*(f(x,\eta z)) &= \log\left(\sum_{i=1}^n \exp(x_i + \log(1-\eta z_i))\right) \\
 &= \log\left(\sum_{i=1}^n \exp(x_i)(1-\eta z_i)\right) \\
 &= \log\left(\sum_{i=1}^n \exp(x_i) - \eta \sum_{i=1}^n \exp(x_i) z_i\right) \\
 &\leq \log\left(\sum_{i=1}^n \exp(x_i)\right) - \eta \frac{\sum_{i=1}^n \exp(x_i) z_i}{\sum_{i=1}^n \exp(x_i)} \\
 &= \psi^*(x) - \eta w^Tz,
\end{align*}
so condition 1 is satisfied.

Also, using the fact that $\log(1-\eta z_i) \geq -\eta z_i - \eta^2 z_i^2$, we have
\begin{align*}
w^T(x-\eta z - f(x, \eta z)) &= w^T(x-\eta z - (x + \log(1-\eta z))) \\
 &\leq \eta^2 \sum_{i=1}^n w_iz_i^2 \\
 &\leq \eta^2 r \sum_{i=1}^n |w_i||z_i|,
\end{align*}
where in the last step we used the fact that $|z_i| \leq r$. These updates 
thus fulfill the conditions of Theorem~\ref{thm:main} for $\alpha$ equal to 
$r$.

\end{document}