### 2014-01-25 lp/algorithms.tex
initial paper

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Algorithms} 
\label{sec:algorithms}
We now state our two main algorithms and their associated 
regret bounds.

\end{document}