### 2014-01-25 lp/paper_icml.tex

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014, times}
\usepackage{tabularx}
\input{latex-defs.tex}

%\allowdisplaybreaks

\icmltitlerunning{Adaptive Exponentiated Gradient}

\begin{document}

\twocolumn[
\icmltitle{Adaptive Exponentiated Gradient}

\icmlauthor{Jacob Steinhardt}{jsteinhardt@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}
\icmlauthor{Jonathan Huggins}{jhuggins@mit.edu}
\icmladdress{MIT CSAIL,
             77 Massachusetts Avenue, Cambridge, MA 02139 USA}
\icmlauthor{Percy Liang}{pliang@cs.stanford.edu}
\icmladdress{Stanford University,
             353 Serra Street, Stanford, CA 94305 USA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{online learning, optimistic, adaptive, matrix multiplicative weights}

\vskip 0.3in
]

\begin{abstract} 
We present an adaptive variant of the exponentiated gradient algorithm, 
achieving regret that depends only on the variance of the best retrospective 
expert. This generalizes and improves a result by 
\cite{hazan2010variation}. Our analysis is conceptually clean and brings together 
two recent developments in online learning, as well as a novel construction 
using an auxiliary instance of follow the regularized leader to tune a regret 
bound. Using the machinery developed, we are also able to extend our results 
to the matrix case, presenting an \emph{adaptive matrix exponentiated gradient} 
algorithm. Our proof in this case involves a novel analysis tool generalizing 
follow the regularized leader to vector-valued payoffs, which may be of 
independent interest.
\end{abstract}

\subfile{introduction.tex}
\subfile{algorithms.tex}

\bibliography{adaptive}
\bibliographystyle{icml2014}

\end{document}