### 2014-01-25 lp/matrix-theorem.tex

\documentclass[11pt]{article}
\usepackage{import}
\input{latex-defs.tex}
\DeclareMathOperator{\Regret}{Regret}
\usepackage{fullpage}

\begin{document}
\section{Proof for Matrix Setting}
We assume that $-I \preceq Z_t \preceq I$  and that $\eta \leq \frac{1}{2}$. We 
consider $W$ bounded so that $W \succeq 0$, $\Tr(W) \leq 1$. Thus when we write 
$W \propto A$ we mean that $W = \frac{A}{\Tr{A}}$.
\begin{theorem}
\label{thm:matrix}
For a sequence $S_1, \ldots, S_T$ of symmetric matrices, consider the following 
prediction algorithm:
\begin{enumerate}
\item Predict $W_t \propto \exp(X_t - \eta S_t)$.
\item Receive loss $\Tr(W_t^{\top}Z_t)$.
\item Update $X_{t+1} = X_t - \eta Z_t - \eta^2 (Z_t-S_t)^2$.
\end{enumerate}
Then the regret relative to a fixed matrix $W$ is bounded by
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr((Z_t-S_t)^{\top}W(Z_t-S_t)). \]
\end{theorem}
% To prove the theorem, we will make use of the following lemma:
% \begin{lemma}
% \label{lem:main}
% Let $\psi(W)$ be a convex function that is bounded below, and 
% let $\psi^*(X) \eqdef \sup_{W} \Tr(W^TX) - \psi(W)$. Let $S$ be any convex set. 
% 
% Suppose that 
% we generate a sequence $X_t$ via $X_0 = 0$ and 
% $X_{t+1} = X_t - \eta Z_t - \eta^2 U_t$, which satisfies the 
% property 
% \[ \psi^*(X_t - \eta Z_t - \eta^2 U_t) \leq \psi^*(X_t) - \eta \partial \psi^*(X_t)^TZ_t. \]
% Then, defining $W_t \eqdef \arg\max_{\tilde{W} \in S} \Tr(\tilde{W}^TX) - \psi(\tilde{W})$, for any $W \in S$ we have 
% the bound
% \[ \sum_{t=1}^T \Tr(W_t^TX_t) - \sum_{t=1}^T \Tr(W^TX_t) \leq \frac{\psi(W) + \psi^*(0)}{\eta} + \eta \sum_{t=1}^T \Tr(W^TU_t). \]
% \end{lemma}

\begin{proof}[Proof of Theorem~\ref{thm:matrix}]
Define $\psi(W) = \Tr(W[\log(W)-I])$ and $\psi^*(X) \eqdef \sup_{W \succeq 0} \Tr(W^{\top}X) - \psi(W) = \Tr(\exp(X))$. 
Note that $\partial \psi^*(X)^{\top}Z = \Tr(\exp(X)Z)$.

We will prove a pair of bounds that will then allow us to obtain a regret bound by telescoping $\psi^*(X_t)$. 
Our first bound is:
\begin{equation}
\label{eqn:bound-1}
\Tr(\exp(X_t - \eta Z_t - \eta^2 (Z_t - S_t)^2 )) \leq \Tr(\exp(X_t - \eta S_t)) - \eta \Tr(\exp(X_t - \eta S_t)(Z_t - S_t)).
\end{equation}
This bound is the crux of our argument (the rest of the manipulations are standard and don't make use of 
any specific properties of our setting other than convexity of $\psi$). We can verify the bound 
using the Golden-Thompson inequality $\Tr(\exp(A+B)) \leq \Tr(\exp(A)\exp(B))$ together 
with the fact that $-X-X^2 \preceq \log(I-X)$ for $-\frac{1}{2}I \preceq X \preceq \frac{1}{2}I$:
\begin{align}
\Tr(\exp(X_t - \eta Z_t - \eta^2 (Z_t - S_t)^2)) &= \Tr(\exp((X_t - \eta S_t) - \eta (Z_t - S_t) - \eta^2 (Z_t - S_t)^2)) \\
 &\leq \Tr(\exp(X_t - \eta S_t)\exp(-\eta (Z_t - S_t) - \eta^2 (Z_t - S_t)^2)) \\
 &\leq \Tr(\exp(X_t - \eta S_t)\exp(\log(I - \eta (Z_t - S_t)))) \\
 &= \Tr(\exp(X_t - \eta S_t)(I - \eta (Z_t - S_t)) \\
 &= \Tr(\exp(X_t - \eta S_t)) - \eta \Tr(\exp(X_t - \eta S_t)(Z_t - S_t)).
\end{align}
Our second bound is
\begin{equation}
\label{eqn:bound-2}
\Tr(\exp(X_t - \eta S_t)) \leq \Tr(\exp(X_t)) - \eta \Tr(\exp(X_t - \eta S_t)S_t).
\end{equation}
This is just saying that $\psi^*(X_t - \eta S_t) \leq \psi^*(X_t) - \eta \langle \partial \psi^*(X_t - \eta S_t), S_t \rangle$, 
which follows directly from convexity of $\psi^*$.

Combining (\ref{eqn:bound-1}) and (\ref{eqn:bound-2}) and defining $\tilde{W}_t \eqdef \exp(X_t - \eta S_t)$ yields
\begin{equation}
\label{eqn:bound-3}
\Tr(\exp(X_{t+1})) \leq \Tr(\exp(X_t)) - \eta \Tr(\tilde{W}_t^{\top}Z_t).
\end{equation}
We thus have
\begin{align}
\sum_{t=1}^T \Tr(\tilde{W}_t^{\top}Z_t) & \leq \frac{\psi^*(0) - \psi^*(X_{T+1})}{\eta} \\
 &\leq \frac{\psi^*(0) + \psi(W) + \sum_{t=1}^T \Tr(W^{\top}(\eta Z_t + \eta^2 (Z_t - S_t)^2))}{\eta} \\
 &\leq \frac{\psi^*(0) + \psi(W)}{\eta} + \sum_{t=1}^T \Tr(W^{\top}Z) + \eta \sum_{t=1}^T \Tr(W^{\top}(Z_t - S_t)^2) \\
 &= \frac{\psi^*(0) + \psi(W)}{\eta} + \sum_{t=1}^T \Tr(W^{\top}Z) + \eta \sum_{t=1}^T \Tr((Z_t-S_t)^{\top}W(Z_t - S_t)).
\end{align}
We thus see that, had we predicted with $\tilde{W}_t$, we would have essentially the desired regret bound. 
However, we instead predicted with $W_t = \frac{\tilde{W}_t}{\Tr(\tilde{W}_t)}$, and so we require one additional step.

This additional step is captured in the following lemma:
%% \begin{lemma}
%% \label{lem:project}
%% Define $\psi_{\sH}^*(X) \eqdef \sup_{W \in \sH} \Tr(W^{\top}X) - \psi(W)$. Then
%% \[ \psi_{\sH}^*(X
(TODO: add in additional step, may not be easily stated as a separate lemma unfortunately)
\end{proof}

\section{FTRL-$\sK$}

In order to fully exploit Theorem~\ref{thm:matrix}, we need a 
generalization of the Follow the Regularized Leader lemma, where 
losses are vector-valued and are ordered by some cone, rather than 
via the usual total ordering on $\bR$. To deal with the cone structure, 
we define the notion of a \emph{global minimizer}:
\begin{definition}
Let $\sK$ be a cone in $\bR^d$ and let $f : \bR^n \to \bR^d$ be a 
function. We call $X$ a \emph{global minimizer} of $f$ if $f(X) \leq_{\sK} f(X')$ 
for all $X'$.
\end{definition}

We prove the following:
\begin{lemma}
\label{lem:ftrl-k}
Suppose that for all $1 \leq t \leq T+1$, there 
exists a global minimizer $M_t$ of 
$\psi(X) + \sum_{s=1}^{t-1} f_s(X)$. Then for all $M$,
\[ \sum_{t=1}^T f_t(M_t) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \]
\end{lemma}
\begin{proof}
We will prove the lemma by induction on $T$. Note that the lemma is equivalent to showing that 
\[ \psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) \leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M) \]
for all $M$.

In the base case $T = 0$, we have
\[ \psi(M_1) \leq_{\sK} \psi(M), \]
which follows from the fact that $M_1$ is a global minimizer of $\psi$ and 
hence $\psi(M_1) \leq_{\sK} \psi(M)$ for all $M$. For the inductive step, suppose that
\[ \sum_{t=1}^{T-1} f_t(M_{t+1}) \leq_{\sK} \psi(M) \sum_{t=1}^{T-1} f_t(M) \]
for all $M$, and invoke this for the particular choice $M = M_{T+1}$. Then we have
\begin{align}
\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) &= \psi(M_1) + \sum_{t=1}^{T-1} f_t(M_{t+1}) + f_T(M_{T+1}) \\
 &\leq_{\sK} \psi(M_{T+1}) + \sum_{t=1}^{T-1} f_t(M_{T+1}) + f_T(M_{T+1}) \\
 &= \psi(M_{T+1}) + \sum_{t=1}^T f_t(M_{T+1}) \\
 &\leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M)
\end{align}
for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
the last inequality. This completes the inductive hypothesis and hence the proof.
\end{proof}

The condition that a global minimizer always exists may seem somewhat strong, so we now invoke it for a concrete case.
\begin{corollary}
\label{cor:ftrl-matrix}
Let $\sK$ be the semidefinite cone, $\psi(M) = M^{\top}M$, and $f_t(M) = (M - Z_t)^{\top}(M - Z_t)$. 
Then the unique global minimizer of $\psi(M) + \sum_{s=1}^{t-1} f_t(M)$ is $M_t \eqdef \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. 
In particular, 
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M_{T+1})^2 + 4\log(T+1)I. \]
\end{corollary}

\begin{proof}
To show that $M_t$ is the global minimizer, consider some other value $\tilde{M}_t = M_t + D_t$. Then we have 
\begin{align}
\lefteqn{\tilde{M}_t^{\top}\tilde{M}_t + \sum_{s=1}^{t-1} (Z_s - \tilde{M}_t)^{\top}(Z_s - \tilde{M}_t)} \\
 &= (M_t + D_t)^{\top}(M_t + D_t) + \sum_{s=1}^{t-1} (Z_s - M_t - D_t)^{\top}(Z_s - M_t - D_t) \\
 &= M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t) + tD_t^{\top}D_t + D_t^{\top}\left[ M_t + \sum_{s=1}^{t-1} (Z_s - M_t) \right] + \left[ M_t + \sum_{s=1}^{t-1} (Z_s - M_t)\right]^{\top} D_t \\
 &= M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t) + tD_t^{\top}D_t \\
 &\succeq M_t^{\top}M_t + \sum_{s=1}^{t-1} (Z_s - M_t)^{\top}(Z_s - M_t).
\end{align}
Since this holds for all $D_t$, $M_t$ is indeed a global minimizer. Thus, by Lemma~\ref{lem:ftrl-k}, we have
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M_{T+1})^2 + \sum_{t=1}^T (Z_t-M_t)^2 - (Z_t-M_{t+1})^2. \]
To complete the corollary, we need only show that $(Z_t-M_t)^2 - (Z_t-M_{t+1})^2 \preceq \frac{4}{t+1}I + M_t^2 - M_{t+1}^2$ 
(since the $M_t^2 - M_{t+1}^2$ terms will telescope to $M_1^2 - M_{T+1}^2 = -M_{T+1}^2 \preceq 0$).
To show this, note that $M_{t+1} - M_t = \frac{1}{t+1}(Z_t - M_t)$ and hence
\begin{align}
(Z_t - M_t)^2 - (Z_t - M_{t+1})^2 &= (M_{t+1}-M_t)Z_t + Z_t(M_{t+1}-M_t) + M_t^2 - M_{t+1}^2 \\
  &= \frac{1}{t+1}\left[(Z_t - M_t)Z_t + Z_t(Z_t-M_t)\right] + M_t^2 - M_{t+1}^2 \\
  &\preceq \frac{4}{t+1}I + M_t^2 - M_{t+1}^2.
\end{align}
This completes the corollary.
\end{proof}

% \begin{corollary}
% If we set $S_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$ (so $S_1 = 0$, $S_2 = \frac{1}{2}Z_1$, $S_3 = \frac{1}{3}(Z_1+Z_2)$, etc.), 
% then we obtain the bound
% \[ \sum_{t=1}^T \Tr(W_t^TZ_t) - \sum_{t=1}^T \Tr(W^TZ_t) \leq \frac{\log(n)+\Tr(W\log(W))}{\eta} + \eta \sum_{t=1}^T \Tr((Z_t-M)^TW(Z_t-M)) + O(1), \]
% where $M \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
% \end{corollary}
% \begin{proof}
% We will show that $\sum_{t=1}^T \Tr((Z_t-S_t)^2) \leq \sum_{t=1}^T \Tr((Z_t-M)^2) + O(1)$. (This follows from the fact that 
% $S_{t} = \argmin_S \Tr(S^2) + \sum_{s=1}^{t-1} \Tr((Z_s-S)^2)$, and hence can be interpreted as Follow the Regularized Leader.)
% 
% Next, we use the fact that, 
% \end{proof}

\end{document}