### 2014-01-25 lp/machinery.tex
done with mw12 draft

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Adaptive Optimistic Learning} 
\label{sec:machinery}

\end{document}