### 2014-01-25 lp/ftrl-aux.tex
done with machinery

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Choosing $m_t$ via Follow the Regularized Leader} 
\label{sec:ftrl-aux}

\end{document}