### 2014-01-25 lp/machinery.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Adaptive Optimistic Learning} 
\label{sec:machinery}
In this section we present a framework for \emph{adaptive optimistic learning}. 
This combines the framework of adaptive learning presented in \cite{orabona2013general} with 
the framework of optimistic learning in \cite{rakhlin2012}. While the combination is 
straightforward, we will show that in our case the combination of the two yields meaningfully 
better results than using either one in isolation. The adaptive optimistic mirror descent 
algorithm thus obtained is given in Algorithm~\ref{alg:adaptive-optimistic}.

\begin{algorithm}
\caption{Adaptive Optimistic Mirror Descent}
\label{alg:adaptive-optimistic}
\begin{algorithmic}
\STATE $\theta_1 = 0$
\FOR{$t=1$ {\bfseries to} $T$}
  \STATE Choose $w_t = \nabla \psi_t^*(\theta_t - \eta m_t)$
  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}

The key regret bound for Algorithm~\ref{alg:adaptive-optimistic} is 
given in Theorem~\ref{thm:adaptive-optimistic}:
\begin{theorem}
Algorithm~\ref{alg:adaptive-optimistic} achieves regret bounded by
\begin{equation}
\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}m_t\right).
\end{equation}
\end{theorem}
\begin{proof}
TODO
\end{proof}
As an immediate consequence, we have
\begin{corollary}
\label{cor:a-o1}
Suppose that $\psi_{t+1}$ is chosen so that $\psi_{t+1}^*(\theta_t - \eta z_t) \leq \psi_t^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. Then
\begin{equation}
\Regret(w) \leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}.
\end{equation}
\end{corollary}
\begin{proof}
TODO
\end{proof}
Finally, we consider a simplified version of Algorithm~\ref{alg:adaptive-optimistic}, 
which corresponds to choosing $\psi_t(w) = \psi(w) + \eta^2 w^{\top} \sum_{s=1}^{t-1} u_s$ 
for some fixed function $\psi$.

\begin{algorithm}
\caption{Adaptive optimistic mirror distance (additive version)}
\label{alg:a-o2}

\begin{algorithmic}
\STATE Initialize $\theta_1$ arbitrarily
\FOR{$t=1$ {\bfseries to} $T$}
  \STATE Choose $w_t = \nabla \psi^*(\theta_t - \eta m_t)$
  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t - \eta^2 u_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}

\begin{corollary}
\label{cor:a-o2}
Suppose that $u_t$ is chosen so that $\psi^*(\theta_t - \eta z_t - \eta^2 u_t) \leq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
Then 
\begin{equation}
\Regret(w) \leq \frac{\psi^*(\theta_1)}{\eta} + \eta w^{\top} \sum_{t=1}^T u_t.
\end{equation}
\end{corollary}
\begin{proof}
TODO
\end{proof}
To give some intuition for the condition in Corollary~\ref{cor:a-o2}, note that 
$w_t = \nabla \psi^*(z_t - \eta m_t)$, and so, since $\psi^*$ is convex, we have
$\psi^*(\theta_t - \eta z_t) \geq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
The term $\eta^2 u_t$ is thus a second-order correction to flip the sign of the inequality 
(by taking a Taylor expansion, one can show that $\eta^2$ is the right order of magnitude for 
the correction term).

\paragraph{Application to exponentiated gradient.} We now show how to apply 
Corollary~\ref{cor:a-o2} to obtain an adaptive exponentiated gradient algorithm.
\begin{proposition}[Adaptive Exponentiated Gradient]
\label{prop:aeg}
Consider the updates given by $\theta_{1,i} = -\log(n)$ and 
$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2$, 
with prediction $w_{t,i} = \exp(\theta_{t,i} - \eta m_{t,i})$. For this 
algorithm,
\begin{equation}
\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
\end{equation}
\end{proposition}
\begin{proof}
TODO
\end{proof}
Note that in the case that $m_{t,i} = 0$, the update is 
$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 z_{t,i}^2$, whereas multiplicative 
updates by $1 - \eta z_{t,i}$ would correspond to 
$\theta_{t+1,i} = \theta_{t,i} + \log(1 - \eta z_{t,i})$. Since 
$-\eta z_{t,i} - \eta^2 z_{t,i}^2 \leq \log(1 - \eta z_{t,i})$ when 
$|\eta z_{t,i}| \leq \frac{1}{2}$, we can think of these updates as a second-order 
under-approximation to (\ref{eqn:mw2}), which however can more naturally incorporate 
the optimistic updates based on $m_t$.

In the next section, we will show how to adaptively choose $m_t$ to optimize the regret bound 
in Proposition~\ref{prop:aeg}.

\end{document}