### 2014-01-26 lp/introduction.tex
fixed missing references

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Introduction} 
\label{sec:intro}
The exponentiated gradient algorithm is a powerful tool for performing online 
learning in the presence of many irrelevant features 
\cite{kivinen1997, littlestone1988}. One common application of exponentiated 
gradient is to the learning from experts setting, in which it is also known 
as the weighted majority algorithm \cite{littlestone1989wm} and entertains regret bounds of the 
form 
\[ \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2, \]
where $w_t$ is the weight vector of the learner at time $t$ and $z_t$ is the 
vector of losses. Such bounds can be obtained under the mirror descent 
framework using a local norms analysis \cite{shalev2011}.

In contrast, \cite{cesa2007} present a variant 
of this algorithm based on a multiplicative update of $(1 - \eta z_{t,i})$ rather 
than the usual update of $\exp(-\eta z_{t,i})$. This algorithm cannot be 
straightforwardly cast in the mirror descent framework, yet achieves an 
improved  regret bound of
\begin{equation}
\label{eqn:mw-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2.
\end{equation}
The difference is that the regret is in terms of the competing expert $w$ instead 
of the learner's strategy $w_t$. This latter bound can sometimes be much 
stronger; we show in Proposition~\ref{prop:separation} that there is a 
$\Theta(\sqrt{T})$ separation in worst-case regret in what we refer to as the 
``quasi-realizable'' setting. The difference between these two types of updates 
is also discussed in \cite{mw-survey}. % TODO: should we actually cite this?

The fact that an algorithm achieving a better regret bound cannot be cast in the 
mirror descent framework is a bit unsettling. Does this mean that we should 
abandon mirror descent as the gold standard for online learning, despite 
theorems asserting its optimality \cite{srebro2011}? We answer this question 
in the negative: the $1-\eta z_{t,i}$ can be understood as an \emph{adaptive 
mirror descent algorithm}: one where the regularizer changes in each round 
in response to previously observed vectors $z_t$. We obtain a 
particularly natural interpretation of the updates as performing a form of 
second-order correction on the gradient step.

By exploiting this connection, we show that the perhaps more natural update 
of $\exp(-\eta z_{t,i} - \eta^2 z_{t,i}^2)$ achieves the same regret bound 
as in (\ref{eqn:mw-regret}), and that both updates continue to enjoy this 
regret even when generalized from the simplex to an arbitrary convex set 
(including all of $\bR^n$, which corresponds to unconstrained exponentiated 
gradient).

Examining (\ref{eqn:mw-regret}) more closely, we see that we should expect 
to perform well when the best expert $i$ has losses consistently close to zero: 
then we set $w_j$ to $\delta_{ij}$ and the second term in (\ref{eqn:mw-regret}) 
becomes $\sum_{t=1}^T z_{t,i}^2 \approx 0$. However, this assumption may be 
unrealistic and a more realistic assumption is that the best expert's losses 
have low variance: $\sum_{t=1}^T (z_{t,i} - m)^2$ is small for some $m$. More 
generally, we would like to obtain a regret bound of the form
\begin{equation}
\label{eqn:variance-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i}-m_i)^2
\end{equation}
with $m$ chosen to be the optimal value of 
$m^* \eqdef \frac{1}{T} \sum_{t=1}^T z_t$.

We achieve such a regret bound by combining two ideas: the first is 
Rakhlin and Sridharan's framework of \emph{optimistic learning} 
\cite{rakhlin2012}, which in our setting yields bounds of the form 
\begin{equation}
\label{eqn:optimistic-regret}
\frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
\end{equation}
If we could set $m_t$ to be $m^*$, then we would achieve the bound in 
(\ref{eqn:variance-regret}). Unfortunately, $m^*$ is only known after observing 
the full sequence of gradients. To overcome this, we learn $m^*$ itself in an 
online fashion, choosing the $m_t$ according to a Follow the Regularized Leader 
(FTRL) algorithm \cite{hazan2011ftrl}. The regret bound thus obtained is given in 
Corollary~\ref{cor:EG-variance}, which improves a result by 
\cite{hazan2010variation}.

Finally, we extend all these results to the matrix setting, where $W_t$ is now 
a positive semidefinite matrix, possibly constrained to have trace $1$ 
(in analogy with the simplex). This setting has already been extensively studied 
\cite{tsuda2005matrix, arora2007combinatorial} and is important in 
obtaining online and approximation bounds for various combinatorial optimization 
problems \cite{arora2007combinatorial, hazan2012near}. Using the machinery so 
far developed, all of our results extend straightforwardly to the matrix setting, 
with the exception of the auxiliary FTRL algorithm for estimating $M^*$. In this 
case we need a new algorithm and analysis tool: a variant of FTRL for 
\emph{vector-valued} losses ordered relative to some cone $\sK$. This generalized 
FTRL lemma is given in Lemma~\ref{lem:ftrl-k} and may be of independent interest.

For convenience, we summarize our contributions below:
\begin{itemize}
\item An interpretation of the multiplicative weights update as 
      exponentiated gradient with an adaptive regularizer.
\item An improved exponentiated gradient algorithm based on adaptive 
      updates to an entropic regularizer.
\item Further improvements to exponentiated gradient based on optimistic 
      predictions and online learning of the mean gradient.
\item An adaptive matrix exponentiated gradient, with similar improvements 
      based on learning the mean gradient.
\item A generalization of Follow the Regularized Leader to vector-valued 
      loss functions.
\end{itemize}

\paragraph{Related work.} Our work contributes to an already rich literature 
on using adaptive regularizers to obtain 
better regret bounds for online learning algorithms. A common setting is 
adaptive learning of a quadratic regularizer, as in the AROW 
\cite{crammer2009arow}, AdaGrad \cite{duchi2011adagrad}, and online 
preconditioning \cite{streeter2010} algorithms. 
Other work includes dimension-free 
exponentiated gradient \cite{orabona2013dimension} (which, however, 
still performs essentially additive updates) and whitened perceptron 
\cite{cesa2005perceptron}. The non-stationary setting has been explored 
in \cite{vaits2013}, and \cite{mcmahan2010adaptive} obtain regret bounds 
relative to an entire family of regularizers. More recently, many of these 
algorithms have been unified into a single framework by \cite{orabona2013general}.
To our knowledge, adaptively regularized exponentiated gradient has not 
been explored, though \cite{hazan2010variation} provide an ad hoc algorithm 
that is similar to one of ours.

Other notions of ``adaptive'' weights have also been explored; though unrelated 
to the current work, we briefly mention them for completeness. The first 
notion, which has occasionally been referred to as adaptivity, pertains to time-varying experts: here the desire is to obtain regret relative to the best \emph{schedule} of experts, 
where the schedule is in some way restricted to drift slowly over time 
\cite{auer1998tracking, kleinberg2010regret, chiang2012online}. 
The other slightly more related notion is an adaptive step size, where the step 
size $\eta$ changes as a function of time (often at a fixed schedule of 
$1/\sqrt{T}$, but sometimes in a more sophisticated fashion). This idea is used 
extensively throughout the literature but a few particularly relevant papers are 
\cite{hazan2007adaptive, orabona2013dimension}.

Finally, we note that the exponentiated gradient algorithm itself has been analyzed 
and applied in various settings. One common setting, using a hinge-like loss, 
is the Winnow Algorithm \cite{littlestone1988}, which has been carefully analyzed in 
\cite{sabato2012learning}.

\end{document}