### 2014-01-26 lp/matrix.tex
matrix.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Extension to Matrices} 
\label{sec:matrix}
We now extend our results to the matrix setting. Here $W$ is a positive semidefinite 
matrix. The analogue to the simplex is constrainint $\Tr(W)$ to be $1$. Instead of a vector 
of losses $z_t$, we have a matrix of losses $Z_t$, which is assumed to be symmetric and to 
have all eigenvalues between $-1$ and $1$. The loss in round $t$ is $\Tr(W_t^{\top}Z_t)$. 
Since all matrices involves are symmetric, we will typically suppress the transpose and 
just write $\Tr(W_tZ_t)$. Note that we can embed the vector setting in the matrix setting 
by restricting to diagonal matrices.

We start with an extension of Proposition~\ref{prop:aeg} to the matrix setting:
\begin{proposition}
\label{prop:maeg}
Consider the updates given by $\Theta_1 = -\log(n)I$ and 
$\Theta_{t+1} = \Theta_t - \eta Z_t - \eta^2 (Z_t - M_t)^2$, 
with prediction $W_t = \exp(\Theta_t - \eta M_t)$. For this algorithm,
\begin{equation}
\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \Tr(W(Z_t-M_t)^2).
\end{equation}
\end{proposition}
\begin{proof}
TODO
\end{proof}
We would now, as in Section~\ref{sec:ftrl-aux}, like to learn $M_t$ adaptively 
to be competitive with the best fixed value $M^*$. However, it isn't even clear 
that such a best fixed value exists (since the optimal value may depend on $W$). 
Lemma~\ref{lem:global-opt} addresses this issue (proved in slightly greater 
generality for later convenience):
\begin{lemma}
\label{lem:global-opt}
For any $\delta \geq 0$, define $M^* \eqdef \frac{1}{T+\delta} \sum_{t=1}^T Z_t$. Then, for 
any $M' \neq M^*$, we have
\[ \delta (M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 \preceq \delta (M')^2 + \sum_{t=1}^T (Z_t-M')^2. \]
\end{lemma}
\begin{proof}
Write $M' = M^* + D$. Then we have
\begin{align}
\lefteqn{\delta(M')^2 + \sum_{t=1}^T (Z_t - M')^2} \\
 &= \delta(M^*+D)^2 + \sum_{t=1}^T (Z_t - M^* - D)^2 \\
 &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t - M^*)^2 \\
\nonumber
 &\phantom{==} + \left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right]D \\
\nonumber
 &\phantom{==} + D\left[\delta M^* + \sum_{t=1}^T M^*-Z_t\right] + TD^2 \\
 &= \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2 + TD^2 \\
 &\succeq \delta(M^*)^2 + \sum_{t=1}^T (Z_t-M^*)^2,
\end{align}
which completes the lemma.
\end{proof}
Setting $\delta$ to $0$, we now see that $\frac{1}{T} \sum_{t=1}^T Z_t$ is the 
optimal value of $M^*$ for any $W \succeq 0$, in analogy with the vector case. We 
remark that Lemma~\ref{lem:global-opt} is almost purely algebraic, and only relies 
on the property that $D^2 \succeq 0$ for any symmetric matrix $D$. We could obtain 
a similar lemma for any partially ordered algebra equipped with an adjoint operator 
$*$ such that $a^{*}a \succeq 0$ for all elements $a$ of the algebra.

Lemma~\ref{lem:global-opt} gives us a target value for the $M_t$, but we cannot 
simply apply the standard Follow the Regularized Leader lemma. This is because 
we need a result of the form
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*) + \epsilon I, \]
which cannot be straightforwardly expressed as a regret bound. We deal with this 
by deriving a generalization of the FTRL algorithm, which we call FTRL-$\sK$. 
This is an algorithm in which losses are vector-valued and we obtain regret relative 
to a partial ordering defined by a cone $\sK$.

An important notion is that of a \emph{global optimum}. For a function $f : X \to \bR^n$ 
and a cone $\sK \subset \bR^n$, we say that $f$ has a global optimum relative to 
$\sK$ if there exists an $x$ such that $f(x) \leq_{\sK} f(y)$ for all $y \in X$.

\begin{lemma}
\label{lem:ftrl-k}
Suppose that for all $1 \leq t \leq T+1$, there 
exists a global minimizer $M_t$ of 
$\psi(X) + \sum_{s=1}^{t-1} f_s(X)$. Then for all $M$,
\[ \sum_{t=1}^T f_t(M_t)-f_t(M) \leq_{\sK} \psi(M) - \psi(M_1) + \sum_{t=1}^T f_t(M_{t}) - f_t(M_{t+1}). \]
\end{lemma}
\begin{proof}
We will prove the lemma by induction on $T$. Note that the lemma is equivalent to showing that 
\[ \psi(M_1) + \sum_{t=1}^T f_t(M_{t+1}) \leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M) \]
for all $M$.

In the base case $T = 0$, we have
\[ \psi(M_1) \leq_{\sK} \psi(M), \]
which follows from the fact that $M_1$ is a global minimizer of $\psi$ and 
hence $\psi(M_1) \leq_{\sK} \psi(M)$ for all $M$. For the inductive step, suppose that
\[ \sum_{t=1}^{T-1} f_t(M_{t+1}) \leq_{\sK} \psi(M) \sum_{t=1}^{T-1} f_t(M) \]
for all $M$, and invoke this for the particular choice $M = M_{T+1}$. Then we have
\begin{align}
\lefteqn{\psi(M_1) + \sum_{t=1}^T f_t(M_{t+1})} \\
 &= \psi(M_1) + \sum_{t=1}^{T-1} f_t(M_{t+1}) + f_T(M_{T+1}) \\
 &\leq_{\sK} \psi(M_{T+1}) + \sum_{t=1}^{T-1} f_t(M_{T+1}) + f_T(M_{T+1}) \\
 &= \psi(M_{T+1}) + \sum_{t=1}^T f_t(M_{T+1}) \\
 &\leq_{\sK} \psi(M) + \sum_{t=1}^T f_t(M)
\end{align}
for all $M$, where we use the fact that $M_{T+1}$ is a global minimizer of $\psi(M) + \sum_{t=1}^T f_t(M)$ for 
the last inequality. This completes the inductive hypothesis and hence the proof.
\end{proof}

As a corollary, we have (compare to Proposition~\ref{prop:ftrl-aux}):
\begin{corollary}
\label{cor:ftrl-k-aux}
Suppose that we choose $M_t = \frac{1}{t} \sum_{s=1}^{t-1} Z_s$. Then
\[ \sum_{t=1}^T (Z_t - M_t)^2 \preceq \sum_{t=1}^T (Z_t - M^*)^2 + 4\log(T+1)I, \]
for $M^* \eqdef \frac{1}{T} \sum_{t=1}^T Z_t$.
\end{corollary}
Combining Proposition~\ref{prop:maeg} with Corollary~\ref{cor:ftrl-k-aux} gives a 
regret bound for the MEG-Variance algorithm, whose updates are given in 
Table~\ref{tab:algorithms}.
\begin{corollary}
\label{cor:maeg-variance}
The MEG-Variance algorithm achieves a regret bound of
\[ \Regret(w) \leq \frac{\log(n)}{\eta} + \eta \left[ \sum_{t=1}^T \Tr(W(Z_t-M^*)^2) + 4\log(T+1) \right]. \]
\end{corollary}

\end{document}