### 2014-01-26 lp/conclusion.tex
conc

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Conclusion} 
\label{sec:conclusion}

\end{document}