### 2014-01-27 lp/machinery.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Adaptive Optimistic Learning} 
\label{sec:machinery}
In this section we present a framework for \emph{adaptive optimistic learning}. 
This combines the framework of adaptive learning presented in \cite{orabona2013general} with 
the framework of optimistic learning in \cite{rakhlin2012}. While the combination is 
straightforward, we will show that in our case the combination of the two yields meaningfully 
better results than using either one in isolation. The adaptive optimistic mirror descent 
algorithm thus obtained is given in Algorithm~\ref{alg:adaptive-optimistic}.

\begin{algorithm}
\caption{Adaptive Optimistic Mirror Descent}
\label{alg:adaptive-optimistic}
\begin{algorithmic}
\STATE $\theta_1 = 0$
\FOR{$t=1$ {\bfseries to} $T$}
  \STATE Choose $w_t = \nabla \psi_t^*(\theta_t - \eta m_t)$
  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}
Note that Algorithm~\ref{alg:adaptive-optimistic} does not contain a 
projection step back onto the feasible domain $S$. If $\psi_t$ is defined to be 
infinite outside of $S$, then $w_t$ will always lie in $S$ and hence the 
projection step is unnecessary. (However, modifying $\psi_t$ in this way may 
complicate the analysis, i.e. by giving $\psi_t^*$ a more complicated form.)

The regret bound for Algorithm~\ref{alg:adaptive-optimistic} is 
given in Theorem~\ref{thm:adaptive-optimistic}. A key observation is that 
the regret is in terms of $z_t - m_t$ rather than $z_t$ (so the closer 
$m_t$ is to $z_t$, the smaller the regret; this will be more apparent 
in various corollaries).
\begin{theorem}
\label{thm:adaptive-optimistic}
Algorithm~\ref{alg:adaptive-optimistic} achieves regret bounded by
\begin{dmath}
\label{eqn:regret-1}
\frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t - \eta m_t) + \eta w_t^{\top}(z_t-m_t)\right).
\end{dmath}
\end{theorem}
\begin{proof}
First note that $\psi_t^*$ is convex and that 
$w_t = \nabla \psi_t^*(\theta_t - \eta m_t)$. It thus follows that 
$\psi_t^*(\theta_t) \geq \psi_t^*(\theta_t - \eta m_t) + \eta m_t^{\top}z_t$. 
Then, by definition of the Fenchel conjugate together with telescoping sums, we 
have, for any $w$, 
\begin{align*}
\lefteqn{\hskip -0.1in w^{\top}\theta_{T+1} - \psi_{T+1}(w)} \\
\hskip 0.1in  &\leq \psi_{T+1}^*(\theta_{T+1}) \\
  &= \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) \\
  &\leq \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t.
\end{align*}
Re-arranging and expanding $\theta_{T+1}$, we have
\begin{dmath}
-\eta w^{\top} \sum_{t=1}^T z_t \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^{T} \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t-\eta m_t) - \eta w_t^{\top}m_t.
\end{dmath}
Adding $\eta \sum_{t=1}^T w_t^{\top}z_t$ to both sides, we obtain
\begin{dmath}
\eta \Regret(w) \leq \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t) + \eta w_t^{\top}(z_t-m_t),
\end{dmath}
which yields the desired result.
\end{proof}
As an immediate consequence, we have
\begin{corollary}
\label{cor:a-o1}
Suppose that $\psi_{t+1}$ is chosen so that $\psi_{t+1}^*(\theta_t - \eta z_t) \leq \psi_t^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. Then
\begin{equation}
\Regret(w) \leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}.
\end{equation}
\end{corollary}
\begin{proof}
Since $\theta_{t+1} = \theta_t - \eta z_t$, if the condition in the corollary 
holds then the sum in the regret bound (\ref{eqn:regret-1}) is bounded above 
term-wise by zero. 
Hence we obtain $\Regret(w) \leq \frac{1}{\eta}\left(\psi_1^*(\theta_1) + \psi_{T+1}(w)\right)$, as was to be shown.
\end{proof}
Finally, we consider a simplified version of Algorithm~\ref{alg:adaptive-optimistic}, 
which corresponds to choosing $\psi_t(w) = \psi(w) + \eta^2 w^{\top} \sum_{s=1}^{t-1} u_s$ 
for some fixed function $\psi$. In this case, we define 
$\psi_S^*(w) \eqdef \sup_{w \in S} w^{\top}x - \psi(w)$, which 
will allow us to more easily handle restrictions on the domain $S$.

\begin{algorithm}
\caption{Adaptive optimistic mirror distance (additive version)}
\label{alg:a-o2}

\begin{algorithmic}
\STATE Initialize $\theta_1$ arbitrarily
\FOR{$t=1$ {\bfseries to} $T$}
  \STATE Choose $w_t = \nabla \psi_S^*(\theta_t - \eta m_t)$
  \STATE Observe $z_t$ and suffer loss $w_t^{\top}z_t$
  \STATE Update $\theta_{t+1} = \theta_t - \eta z_t - \eta^2 u_t$
\ENDFOR
\end{algorithmic}
\end{algorithm}

\begin{corollary}
\label{cor:a-o2}
Suppose that $u_t$ is chosen so that $\psi^*(\theta_t - \eta z_t - \eta^2 u_t) \leq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
Then 
\begin{equation}
\Regret(w) \leq \frac{\psi_S^*(\theta_1)}{\eta} + \eta w^{\top} \sum_{t=1}^T u_t.
\end{equation}
\end{corollary}
Note that, importantly, the conditions of Corollary~\ref{cor:a-o2} are in terms 
of $\psi^*$, while the regret bound and the choice of $w_t$ are in terms of 
$\psi_S^*$. This means that, once we construct an algorithm for the 
unconstrained case, we then achieve a regret bound under arbitrary convex 
constraints.
\begin{proof}
TODO
\end{proof}
To give some intuition for the condition in Corollary~\ref{cor:a-o2}, note that 
$w_t = \nabla \psi^*(z_t - \eta m_t)$, and so, since $\psi^*$ is convex, we have
$\psi^*(\theta_t - \eta z_t) \geq \psi^*(\theta_t - \eta m_t) - \eta w_t^{\top}(z_t - m_t)$. 
The term $\eta^2 u_t$ is thus a second-order correction to flip the sign of the inequality 
(by taking a Taylor expansion, one can show that $\eta^2$ is the right order of magnitude for 
the correction term).

\paragraph{Application to exponentiated gradient.} We now show how to apply 
Corollary~\ref{cor:a-o2} to obtain an adaptive exponentiated gradient algorithm. 
We will state it here in the case of the simplex, since the bounds in this 
case are particularly easy to interpret. Note, however, that our analysis 
goes via Corollary~\ref{cor:a-o2}, and hence applies to arbitrary convex 
constraints.
\begin{proposition}[Adaptive Exponentiated Gradient]
\label{prop:aeg}
Consider the updates given by $\theta_{1,i} = 0$ and 
$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2$, 
with prediction $w_{t,i} \propto \exp(\theta_{t,i} - \eta m_{t,i})$. Then, 
assuming $\|z_t\|_{\infty} \leq 1$, $\|m_t\|_{\infty} \leq 1$ and 
$0 < \eta \leq \frac{1}{4}$, we have
\begin{equation}
\Regret(w) \leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T (z_{t,i} - m_{t,i})^2.
\end{equation}
\end{proposition}
\begin{proof}
We note that, for $S$ equal to the simplex $\Delta_n$ and 
$\psi(w) = \sum_{i=1}^n w_i\log(w_i)$, the $w_t$ is equal to 
$\nabla \psi_S^*(\theta_t - \eta m_t)$. It therefore suffices 
to check that the condition of Corollary~\ref{cor:a-o2} is satisfied with 
$u_{t,i} = (z_{t,i}-m_{t,i})^2$. 
Note that $\psi^*(\theta) = \sum_{i=1}^n \exp(\theta_i)$. We then have
\begin{dgroup*}
\begin{dmath*} \lefteqn{\psi^*(\theta_{t} - \eta z_t - \eta^2 u_t)} 
\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta z_{t,i} - \eta^2 (z_{t,i} - m_{t,i})^2) 
\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i})\exp(-\eta (z_{t,i} - m_{t,i}) - \eta^2 (z_{t,i} - m_{t,i})^2) 
\end{dmath*}\begin{dmath*}  \leq \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i})(1 - \eta (z_{t,i} - m_{t,i})) 
\end{dmath*}\begin{dmath*}  = \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i}) - \eta \sum_{i=1}^n \exp(\theta_{t,i} - \eta m_{t,i}) (z_{t,i} - m_{t,i}) 
\end{dmath*}\begin{dmath*}  = \psi^*(\theta_t - \eta m_t) - \eta \nabla \psi^*(\theta_t - \eta m_t)^{\top} (z_t - m_t).
\end{dmath*}
\end{dgroup*}
The one inequality we made use of was $\exp(-x-x^2) \leq 1-x$ for $|x| < \frac{1}{2}$.
This verifies the condition of Corollary~\ref{cor:a-o2}, yielding a regret bound 
of $\frac{\psi_S^*(0) + \psi(w)}{\eta} + \eta \sum_{i=1}^n w^{\top} u_t$. 
Finally, we note that $\psi_S^*(0) = \log(n)$, 
$\psi(w) = \sum_{i=1}^n w_i\log(w_i) \leq 0$, and 
$u_{t,i} = (z_{t,i} - m_{t,i})^2$, which completes the proof.
\end{proof}
Note that in the case that $m_{t,i} = 0$, the update is 
$\theta_{t+1,i} = \theta_{t,i} - \eta z_{t,i} - \eta^2 z_{t,i}^2$, whereas multiplicative 
updates by $1 - \eta z_{t,i}$ would correspond to 
$\theta_{t+1,i} = \theta_{t,i} + \log(1 - \eta z_{t,i})$. Since 
$-\eta z_{t,i} - \eta^2 z_{t,i}^2 \leq \log(1 - \eta z_{t,i})$ when 
$|\eta z_{t,i}| \leq \frac{1}{2}$, we can think of these updates as a second-order 
under-approximation to (\ref{eqn:mw2}), which however can more naturally incorporate 
the optimistic updates based on $m_t$.

In the next section, we will show how to adaptively choose $m_t$ to optimize the regret bound 
in Proposition~\ref{prop:aeg}.

\end{document}