### 2014-01-27 lp/mw12.tex

\documentclass[paper_icml.tex]{subfiles}
\begin{document}

\section{Two Types of Updates} 
\label{sec:mw12}
Our point of departure is the two different types of multiplicative updates 
mentioned in the introduction. For simplicity we will consider the setting of 
learning from experts, although our later theorems will hold in greater 
generality. In this setting there are $n$ experts, and the learner maintains a 
probability distribution $w_t \in \Delta_n$ over the experts, which gets 
updated in each of $T$ rounds. After selecting the distribution $w_t$, 
the loss vector $w_t \in [-1,1]^n$ of each of the experts is revealed, and the 
learner incurs loss $w_t^{\top}z_t$. The learner's goal is to minimize 
the regret $\sup_{w \in \Delta_n} \Regret(w)$, where
\[ \Regret(w) \eqdef \sum_{t=1}^T w_t^{\top}z_t - \sum_{t=1}^T w^{\top}z_t. \]
We consider two types of updates for the weight vector $w_t$. In both cases 
$w_1$ is initialized as $w_{1,i} = \frac{1}{n}$ for $1 \leq i \leq n$, but 
the updates differ, as in (\ref{eqn:mw1}) and (\ref{eqn:mw2}) below:
\begin{align}
\label{eqn:mw1}
w_{t+1,i} &\propto w_{t,i}\exp(-\eta z_{t,i}) \tag{MW1}\\
\label{eqn:mw2}
w_{t+1,i} &\propto w_{t,i}(1-\eta z_{t,i}) \tag{MW2}
\end{align}
The regret bounds for each of (\ref{eqn:mw1}) and (\ref{eqn:mw2}) are 
well-known \cite{shalev2011, cesa2007}, but we include them below for completeness.
\begin{theorem}
\label{thm:mw12}
For any $0 < \eta \leq \frac{1}{2}$, the updates (\ref{eqn:mw1}) and 
(\ref{eqn:mw2}) enjoy regret, respectively,
\begin{align}
\label{eqn:mw1-regret}
\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n \sum_{t=1}^T w_{t,i}z_{t,i}^2 %\tag{MW1-regret}
 \\
\label{eqn:mw2-regret}
\Regret(w) &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2 %\tag{MW2-regret}
\end{align}
\end{theorem}
To understand why (\ref{eqn:mw2-regret}) may be a better bound than 
(\ref{eqn:mw1-regret}), suppose that the best expert has a loss identically 
equal to zero.\footnote{Of course, if we knew that this was the case ahead of 
time, there would be far better algorithms; we use this scenario purely for 
illustrative purposes.} Then the optimal $w$ places all mass on this expert, 
and (\ref{eqn:mw2-regret}) reduces to $\frac{\log(n)}{\eta} = 2\log(n)$ for 
$\eta = \frac{1}{2}$.

More formally, define a sequence of losses $z_t$ to be \emph{quasi-realizable} 
if one of the experts has identically zero loss and all other experts 
have non-negative cumulative loss, i.e. $\sum_{t=1}^T z_{t,i} \geq 0$. It is 
apparent by the preceding paragraph that (\ref{eqn:mw2}) achieves asymptotically 
constant (as a function of $T$) regret for any quasi-realizable sequence. In 
contrast, for \emph{any} step size $\eta$, there exists a quasi-realizable 
sequence such that (\ref{eqn:mw1}) suffers regret $\Omega(\sqrt{T})$:
\begin{proposition}
\label{prop:separation}
If the updates (\ref{eqn:mw1}) are used with a step size of $\eta$, then for 
any $T$ there exists a quasi-realizable sequence of losses and a vector $w$ 
such that $\Regret(w) = \Omega(\sqrt{T})$.
\end{proposition}
This establishes that the apparent separation between (\ref{eqn:mw1}) and 
(\ref{eqn:mw2}) is real and not an artifact of the analysis. We remark that 
this separation does \emph{not} exist in the realizable case (when all losses 
are non-negative). In this case both (\ref{eqn:mw1}) and (\ref{eqn:mw2}) enjoy 
$O(1)$ regret as a function of $T$.

Finally, note that (\ref{eqn:mw2}) cannot be realized as mirror descent for any 
fixed regularizer. This is because, for any mirror descent algorithm, the 
prediction at time $t+1$ must be a function of $\sum_{t=1}^T z_t$, which is 
not the case for (\ref{eqn:mw2}). Instead, we obtain (\ref{eqn:mw2}) in terms of 
an \emph{adaptive regularizer} $\psi_t(w)$. The mirror descent updates for 
such a regularizer are given by 
\[ w_{t} = \nabla \psi_t^*\left(-\eta \sum_{s=1}^{t-1} z_s\right). \]
See \cite{orabona2013general} for a more complete exposition. For convenience, 
define $\theta_t \eqdef -\eta \sum_{s=1}^{t-1} \theta_s$. The important 
observation in our context is that, by definition of $\psi_{T+1}^*$, 
\begin{align}
- \eta w^{\top} \sum_{t=1}^T z_t &\leq \psi_{T+1}(w) + \psi_{T+1}^*(\theta_{T+1}) \\
\label{eqn:telescope}
 &= \psi_{T+1}(w) + \psi_1^*(\theta_1) + \sum_{t=1}^T \psi_{t+1}^*(\theta_{t+1}) - \psi_t^*(\theta_t).
\end{align}
Motivated by this, we note that if we could choose $\psi_{t+1}$ so that 
$\psi_{t+1}^*(\theta_{t+1}) \leq \psi_{t}^*(\theta_t) - \eta w_t^{\top}z_t$, 
then re-arranging would yield a regret bound of $\frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}$.
\begin{proposition}
\label{prop:correspondence}
Define $\lambda_{t,i} \eqdef \sum_{s=1}^{t-1} \log(1-\eta z_{s,i})$ and let
\[ \psi_t(w) \eqdef \sum_{i=1}^n w_i\log(w_i) + (\theta_t - \lambda_t)^{\top}w. \]
Then adaptive mirror descent with regularizer $\psi_t$ corresponds exactly to 
the updates (\ref{eqn:mw2}), and 
$\psi_{t+1}^*(\theta_{t+1}) \leq \psi_t^*(\theta_t) - \eta w_t^{\top}z_t$. 
The corresponding regret bound is
\begin{align}
\Regret(w) &\leq \frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta} \\
% &= \frac{\log(n)+\sum_{i=1}^n w_i\log(w_i)}{\eta} + \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i} + \frac{\log(1-\eta z_{t,i})}{\eta} \\
 &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2.
\end{align}
\end{proposition}
We will be able to prove this more easily with the machinery developed in 
Section~\ref{sec:machinery}, but we prove it here directly.
\begin{proof}
By standard properties of Fenchel conjugates, we have
\begin{align}
\nabla \psi_t^*(\theta_t) &= \argmin_{w \in \Delta_n} \psi_t(w) - \theta_t^{\top}w \\
 &= \argmin_{w \in \Delta_n} \sum_{i=1}^n w_i\log(w_i) - \lambda_t^{\top}w.
\end{align}
From here we see that 
$w_{t,i} \propto \exp(\lambda_{t,i}) = \prod_{s=1}^{t-1} (1-\eta z_{s,i})$, 
so that $w_{t,i}$ does indeed correspond to the update in (\ref{eqn:mw2}).

Next, we exploit the property that the Fenchel conjugate of $\psi(w) + c^{\top}w$ 
is $\psi^*(\theta - c)$, together with the fact that the Fenchel conjugate of 
$\psi_1(w) = \sum_{i=1}^n w_i\log(w_i)$ is $\psi_1^*(\theta) = \log\left(\sum_{i=1}^n \exp(\theta_i)\right)$, 
to obtain
%\begin{small}
\begin{align*}
\psi_{t+1}^*(\theta_{t+1}) &= \psi_1^*(\theta_{t+1}-(\theta_{t+1}-\lambda_{t+1})) \\
 &= \psi_1^*(\lambda_{t+1}) \\
 &= \log(\sum_{i=1}^n \exp(\lambda_{t+1,i})) \\
 &= \log(\sum_{i=1}^n \exp(\lambda_{t,i} + \log(1 - \eta z_{t,i}))) \\
 &= \log(\sum_{i=1}^n \exp(\lambda_{t,i})(1 - \eta z_{t,i})) \\
 &= \log(\sum_{i=1}^n \exp(\lambda_{t,i}) - \eta \sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}) \\
 &\leq \log(\sum_{i=1}^n \exp(\lambda_{t,i})) - \eta \frac{\sum_{i=1}^n \exp(\lambda_{t,i})z_{t,i}}{\sum_{i=1}^n \exp(\lambda_{t,i})} \\
 &= \psi_t^*(\theta_t) - \eta w_t^{\top}z_t.
\end{align*}
%\end{small}
Thus by the observations following (\ref{eqn:telescope}), $\Regret(w)$ is bounded by 
$\frac{\psi_1^*(\theta_1) + \psi_{T+1}(w)}{\eta}$. Now using the fact that 
$-\eta z - \log(1-\eta z) \leq \eta^2 z^2$ when $|\eta z| \leq \frac{1}{2}$, and also the 
fact that $\sum_i w_i\log(w_i) \leq 0$, we have
\begin{align*}
\lefteqn{\hskip -0.1in \eta\Regret(w)} \\
\hskip 0.1in &\leq \psi_1^*(\theta_1) + \psi_{T+1}(w) \\
 &= \log(n) + \sum_{i=1}^n w_i\log(w_i) %\nonumber \\ &\phantom{==}
    + (\theta_{T+1}-\lambda_{T+1})^{\top}w \\
 &\leq \frac{\log(n)}{\eta} + \eta \sum_{i=1}^n w_i \sum_{t=1}^T z_{t,i}^2,
\end{align*}
which completes the proposition.
\end{proof}
In the next section, we will present a more general framework for \emph{adaptive updates}, 
also incorporating \emph{optimistic prediction} \cite{rakhlin2012}, which will both simplify 
subsequent proofs and also allow us to obtain better bounds.
\end{document}