### 2014-02-05 arXiv/boosted-mirror-descent.tex
created new arXiv version and updated notation

\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Boosted Mirror Descent}
\label{sec:algorithm}

Our results primarily rely on a pair of novel convex optimization algorithms we call primal and dual boosted mirror descent. They are inspired by the decomposition used in \cite{Bach:2012b}.

\subsection{Convex Optimization}

Before introducing the boosted mirror descent (\bmd) algorithms, we establish notation and recall some basic definitions related to convex optimization.

Let $f$ be a real function whose domain is a normed inner product space. The {\em convex conjugate} $f^{*}$ of $f$ is defined as
\[
f^{*}(x') \eqdef \max_{x} \ip{x}{x'} - f(x).
\]
If $f$ is proper, lower semi-continuous, and convex, then by the 
Fenchel-Moreau theorem $f = f^{**}$, i.e.
\[
f(x) = \max_{x'} \ip{x}{x'} - f^{*}(x').
\]
Throughout the paper, we will assume without further notice that functions are proper and lower semi-continuous when required. The (sub)gradients of $f^{*}$ and $f$ can be defined as any argument that maximizes the previous two equations:
\[
\partial f^{*}(x') &\in \argmax_{x} \ip{x}{x'} - f(x) \\
\partial f(x) &\in \argmax_{x'} \ip{x}{x'} - f^{*}(x').
\]

In order to prove fast convergence, we will require additional assumptions beyond convexity. 
Define the \emph{Bregman divergence} of a convex function to be
\[
D_f(x \| x') = f(x)-\ip{\partial f(x')}{x'-x} - f(x'), 
\]
which is a measure of ``how convex'' $f$ is between $x$ and $x'$. 
We say that a function $f$ is \emph{strongly convex with respect 
to a norm $\|\cdot\|$} if
\[
D_f(x \| x') \geq \frac{1}{2}\|x-x'\|^2
\]
A related notion is \emph{strong smoothness}: a convex function $f$ 
is \emph{strongly smooth with respect to $\|\cdot\|$} if
\[ D_f(x \| x') \leq \frac{1}{2}\|x-x'\|^2 \]
An important result is that $f$ is strongly convex with respect to $\|\cdot\|$ 
if and only if its Fenchel conjugate $f^*$ is strongly smooth with respect 
to the dual norm $\|\cdot\|_*$.

\subsection{The Boosted Mirror Descent Algorithms}

We can now define the \bmd  algorithms. Let $X$ and $Y$ be convex subsets 
of the Banach spaces $\mcX$ and $\mcY$, respectively, and let 
$\blf{\cdot}{\cdot}$ be a bilinear map on $\mcX \times \mcY$; typically 
$\blf{\cdot}{\cdot}$ will correspond to an inner product.
% but we will use the bracket notation for convenience even when this is not the case.
Consider the two-argument loss function $L : X \times Y \to \reals$ 
defined as
\[
L(x,y) = \psi(x) + \blf{x}{y} - R(y)
\]
for which we wish to find arguments that obtain $L_{*} = \min_{x}\max_{y} L(x, y)$. Taking only the maximum, we have the one-argument loss function
\[
L(x)
&= \psi(x) + \max_{y \in \Theta} \left\{\blf{x}{y} - R(y) \right\} \\
&= \psi(x) + R^{*}(x).
\]
We can think of $\psi$ as a \emph{primal regularizer} and $R$ as a \emph{dual regularizer}.  The pair of regularization functions $\psi$ and $R$ define a flexible class of functions to optimize, as seen from the following examples:
\begin{enumerate}
\item Let $X$ be the probability simplex over a space $\sS$, let $\phi : \sS \to \reals^d$ be a collection 
      of statistics, and let $Y$ be the unit $L^1$ ball in $\reals^d$. Let $\psi(x) = 0$, 
      $\blf{x}{y} = y^T\bE_{v \sim x}[\phi(v)]$, and 
      $R(y) = \blf{x_0}{y}$. Then $L(x)$ is the 
      \emph{maximum mean discrepancy} between $x$ and $x_0$ relative to $\phi$.
\item \NA{JHH: I don't remember what this example what supposed to be}
	  Let $\psi(x) = 0$ and $R(y) = S^*(i^{-1}(y))$ where $S^*$ is the 
      convex conjugate of any strongly convex function $S : U \to \bR$. Then $L(x) = S(x)$ 
      as long as $i^{-1}(\Theta) = \sB_{U}$.
\item Let $X = Y = \reals^d$. Let $\blf{x}{y} = \ip{x}{y}$, $\psi(x) = \|x\|_1$, $R(x) = \blf{x_0}{y} + \frac{1}{2} \|y\|_2^2$. 
      Then $L(x) = \|x\|_1 + \frac{1}{2} \|x-x_0\|_2^2$. 
\end{enumerate}

We will assume 
throughout that $\psi$ and $R$ are both convex functions, and furthermore that 
$\arg\max_{x} \psi(x) + \blf{x}{y}$ can be efficiently 
computed for all values of $y$.

%In this paper we will consider a family of methods for minimizing $L(x)$, 
%which are based on Bach's generalization of the Frank-Wolfe algorithm and can 
%be interpreted as boosted mirror descent.

Mirror descent, combined with boosting, yields an algorithm for finding 
``saddle points'' of $L(x, y)$. 
For notational convenience, for a sequence of weights $\alpha_1,\alpha_2,\ldots$ 
let 
\(
\bar{x}_t = \frac{\sum_{s=1}^t \alpha_{s}x_{s}}{A_{t}} 
\quad \text{and} \quad 
\bar{y}_t = \frac{\sum_{s=1}^t \alpha_{s}y_{s}}{A_{t}},
\quad \text{where} \quad 
A_{t} \defined \sum_{s=1}^t \alpha_s.
\)

Then 
the {\em primal boosted mirror descent} (\primal) algorithm is:
\begin{enumerate}
\item $x_1 \in \arg\min_x \psi(x)$
\item $y_{t} \in \arg\max_{y \in \Theta} \blf{y}{x_t} - R(y) = \partial R^{*}(x_{t})$
\item $x_{t+1} \in \arg\min_{x} \psi(x) + \blf{x}{\bar{y}_t} = \partial \psi^{*}(-\bar y_{t})$
\end{enumerate}
As long as $\psi$ is strongly convex, for the proper choice of $\alpha_{t}$ we obtain the 
bound (see Corollary~\ref{cor:method-1}):
\begin{equation}
\sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T).
\end{equation}
In other words, $\bar{x}_T$ is close to being a global minimum of $L(x)$.
However, it is often the case that $\psi$ is not strongly convex but $R$ is strongly convex. As we shall see, this will be the case for our application to herding.
In this case we may wish to use the following slightly different algorithm, which we call {\em dual boosted mirror descent} (\dual):
\begin{enumerate}
\item $y_1 \in \arg\max_{y} R(y)$
\item $x_t \in \arg\min_{x} \psi(x) + \blf{x}{y_t} = \partial \psi^{*}(-y_{t})$
\item $y_{t+1} \in \arg\max_{y \in \Theta} \langle y, \bar{x}_t \rangle - R(y) = \partial R^{*}(\bar x_{t})$
\end{enumerate}
We then obtain the following bound 
when $R$ is strongly convex (see Corollary~\ref{cor:method-2}):
\[ \sup_{y \in \Theta} L(\bar{x}_T, y) \leq L_{*} + O(1/T). \]

\subsection{Connections to Other Gradient Descent Algorithms}

As suggested by their names, the \bmd algorithms are closely related: they are dual to each other in the sense that 
performing \primal on $L(x,y)$ is the same as performing \dual on $L'(y, x) \eqdef -L(x,y)$. This relationship is obvious by inspection of the update equations:
\(
\text{Primal} && \text{Dual} \\
y_{t} &= \partial R^{*}(x_{t}) & y_{t+1} &= \partial R^{*}(\bar x_{t}) \\
x_{t+1} &= \partial \psi^{*}(-\bar y_{t}) & x_t &= \partial \psi^{*}(-y_{t})
\)
This duality is analogous to that between mirror descent and conjugate gradient descent \citep{Bach:2012b}. \md minimizes $g(x) + f(x)$ and requires that $g$ be $\alpha$-strongly convex while \cgd minimizes $f^{*}(x) + g^{*}(x)$ and requires that $g$ be $\alpha$-strongly convex. The update equations for \md are
\(
y_{t} &= \partial f(x_{t})  \\
\bar y_{t} &=  (1 - \rho_{t+1})\partial g(x_{t}) - \rho_{t+1}y_{t}  \\
x_{t+1} &=  \partial g^{*}(\bar y_{t}) 
\)
and for \cgd are
\(
 y_{t+1} &= \partial f(x_{t}) \\
\bar y_{t+1} &= (1 - \rho_{t})\bar y_{t} + \rho_{t} y_{t+1}  \\
x_{t} &= \partial g^{*}(-\bar y_{t}).
\)
In fact, the \bmd algorithms are very closely related to \cgd and \md: \cgd can be obtained from \dual and \md obtained from \primal. The relationships are given in Table~\ref{tab:connections}. The table also shows how $\alpha_{t}$-weighted herding and $q$-herding --- which are discussed in the next section --- are obtained from \bmd.

\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{table}[t]
\begin{adjustwidth}{-.5in}{-.5in}
\caption{Formulation of herding and convex optimization algorithms in terms of \bmd. Full details relating \md and \cgd to \bmd are given in the Supplementary Material. See also \citet{Bach:2012b}.}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\begin{tabularx}{1.15\textwidth}{L|lllcL}
Algorithm 			& $\psi(x)$ 	& $R(y)$ & $\alpha_{t}$ & \bmd Algorithm & Notes 
\\ \hline
herding 			& 0			& $\blf{\bar x}{y} + \frac{1}{2} \| y \|_{\sH}^{2}$ 	& 1				 	& \dual & 
\\
$\alpha_{t}$-weighted herding 	& 0	& $\blf{\bar x}{y} + \frac{1}{2} \| y \|_{\sH}^{2}$ 	& unconstrained 	& \dual & $\sum_{s =1}^{t} \alpha_{t} > 0~\forall t$
\\ 
$q$-herding			& 0			& $\blf{\bar\phi}{y} + \frac{1}{q} \| y \|_{q}^{q}$ 	& unconstrained 	& \dual & $1 \le q \le 2$
\\ 
\md					& $g(x)$	& $f^{*}(y)$				& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \primal & 
\\ 
\cgd				& $f^{*}(y)$& $g(x)$					& $\rho_{t}/\prod_{s=1}^{t} (1 - \rho_{s})$				& \dual &
\\
\end{tabularx}
\label{tab:connections}
\end{adjustwidth}
\end{table}
\end{center}
\end{document}