### 2014-02-05 arXiv/chen-proofs.tex

\documentclass[paper.tex]{subfiles}

\begin{document}
\section{An Upper Bound in Infinite Dimensions}
\label{sec:chen-proofs}

An intriguing aspect of classical herding is that it is known to have a convergence rate of 
$O(1/T^2)$ under suitable conditions, as was shown in \citet{Chen:2010a}. In this section, 
we will need to assume that $\psi(x)$ is the indicator function for $X$ (i.e. $0$ on $X$ and 
$\infty$ outside of it) and that $R(y) = \ip{x_0}{y} + R_0(y)$, where 
$R_0$ exhibits \emph{quadratic scaling}: $R_0(sy) = s^2R_0(y)$. Note that 
this implies that $R_0^*(x-x_0) = R^*(x)$.
For simplicity 
we will focus on the case $\alpha_t = 1$. Then we have 
the following bound:
\begin{theorem}
\label{thm:chen}
Suppose that $R$ is strongly convex with respect to the norm $\|\cdot\|$ and 
let $r_* \eqdef \sup_{x \in U} \|x\|_*$. Also suppose that $\psi \equiv 0$ and 
that $R_0$ exhibits quadratic scaling as above. Define 
\[ \gamma^* \eqdef \inf_{y \in y} \sup_{x \in U} \frac{\ip{-y}{x-x_0}}{\sqrt{R(y)}}. \]
Then $R(y_{t+1}) \leq \frac{2r_*^2+(2r_*^2/\gamma_*)^2}{t^2}$ as long 
as $\gamma^* > 0$.
\end{theorem}
This corresponds to Proposition 1 of \citet{Chen:2010b} for $R_{0}(y) = \frac{1}{2}\|y\|_2^2$. Note that if there is an $\ell^2$ ball of radius $\epsilon$ 
around $x_0$ that is contained in $u$ then $\gamma^* \geq \frac{\epsilon}{r_*}$ 
in this case, which is how Chen's result follows from ours.
\begin{proof}[Proof of Theorem~\ref{thm:chen}]
The key quantity we will analyze is $(t-1)^2R(y_t)$. Note that we have
\begin{align*}
\lefteqn{(t-1)^2R(y_t) - t^2R(y_{t+1})} \\
 &= (t-1)^2R^*(\bar{x}_{t-1})-t^2R^*(\bar{x}_t) \\
 &= (t-1)^2R^*(\bar{x}_{t-1})-t^2R_0^*(\bar{x}_t-x_{0}) \\
 &= (t-1)^2\left[R^*(\bar{x}_{t-1})-R_0^*((t/(t-1))(\bar{x}_t-x_{0}))\right] \\
 &= (t-1)^2\left[R^*(\bar{x}_{t-1})-R^*\left(\frac{t\bar{x}_t-x_{0}}{t-1}\right)\right] \\
 &= (t-1)^2\left[\ip{\partial R^*(\bar{x}_{t-1})}{\bar{x}_{t-1}-\frac{t\bar{x}_t-x_{0}}{t-1}}-D_{R^*}\left(\frac{t\bar{x}_t-x_{0}}{t-1} \middle\| \bar{x}_{t-1}\right)\right] \\
 &\geq (t-1)^2\left[\ip{y_t}{-\frac{1}{t-1}(x_t-x_{0})}-\frac{1}{2}\left\|\frac{x_t-x_{0}}{t-1}\right\|_*^2\right] \\
 &= -(t-1)\ip{y_t}{x_t-x_{0}}-\frac{1}{2}\|x_t-x_{0}\|_*^2 \\
 &= -(t-1)\left[\inf_{x \in U} \ip{y_t}{x} - \ip{y_t}{x_{0}}\right]-\frac{1}{2}\|x_t-x_{0}\|_*^2 \\
 &\geq (t-1)\sqrt{R(y_t)}\left[\inf_{y} \sup_{x \in U} \frac{\ip{-y}{x-x_0}}{\sqrt{R(y_t)}}\right]-2r_*^2 \\
 &= \sqrt{(t-1)^2R(y_t)}\gamma^*-2r_*^2.
\end{align*}
There are now two possibilities for each $t$:
\begin{enumerate}
\item $t^2R(y_{t+1}) < (t-1)^2R(y_t)$, or
\item $(t-1)^2R(y_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
\end{enumerate}
In particular, consider the value of $t$ for which 
$t^2R(y_{t+1})$ is maximized. Then by assumption 
case $1$ cannot hold and so $(t-1)^2R(y_t) \leq \frac{4r_*^4}{(\gamma^*)^2}$.
But then by applying the above sequence of inequalities 
we also have 
\begin{align*}
\lefteqn{t^2R(y_{t+1}) - (t-1)^2R(y_t)} \\
 &\leq 2r_*^2 - \sqrt{(t-1)^2R(y_t)}\gamma^* \\
 &\leq 2r_*^2,
\end{align*}
hence $t^2R(y_{t+1}) \leq 2r_*^2 + \frac{4r_*^4}{(\gamma_*)^2}$. 
Since $t^2R(y_{t+1})$ was maximal by assumption, we have
$R(y_{t+1}) \leq \frac{2r_*^2+(2r_*/\gamma_*)^2}{t^2}$, which 
completes the proof.
\end{proof}

\end{document}